---
title: "Voting for the Radical Right by Education"
subtitle: "ESS Round 11 — Bayesian analysis with survey weights"
author:
  - name: Armin Schäfer
    affiliation:
      - id: jgu
        name: Johannes Gutenberg University Mainz
format:
  html:
    number-sections: true
    code-fold: true
    code-summary: "Show the code"
    self-contained: true
    toc: true
execute:
  warning: false
  message: false
---

# Purpose and modelling strategy

This document repeats the class-voting analysis using **education** rather than class as the predictor, to allow a direct comparison between the two stratification variables. It estimates two closely related quantities:

1. **Loyalty:**
$$
   P(\text{radical right}\mid\text{high education})
$$
2. **Contribution:**
$$
\begin{aligned}
   P(\text{high education}\mid\text{radical right}).
\end{aligned}
$$

These are different conditional probabilities. They should not be treated as interchangeable. They are linked by the same underlying $2\times2$ table, so the preferred strategy is to estimate the table-generating probabilities jointly and derive both quantities from the same posterior draws.

The education predictor is `eduhigh`, a dummy variable derived from `edulowhigh` (itself built from ESS `eisced`): `eduhigh = 1` for respondents who have completed at least upper secondary education (`edulowhigh == "high"`), and `eduhigh = 0` for respondents who have not (`edulowhigh == "low"`). **Low education is the reference category** throughout, exactly mirroring how "non-worker" served as the reference category in the class analysis.

The analysis proceeds from simple to more complex models:

- **Model 0:** pooled logistic model, radical-right vote predicted by education.
- **Model 1:** varying country intercepts.
- **Model 2:** varying country intercepts and education slopes.
- **Model 3:** add individual-level covariates.
- **Model 4:** add theoretically motivated interactions with education.

The primary analysis uses the ESS **analysis weight (`anweight`)**. The ESS recommends `anweight` as the default weight for ESS analyses. It combines the post-stratification weight with the population-size adjustment and is therefore appropriate when several countries are pooled. `pspwght` is appropriate for single-country analyses and cross-country comparisons without treating the countries as one population. The present analysis pools ten countries, so `anweight` is the natural default. [ESS weighting guidance](https://www.europeansocialsurvey.org/methodology/ess-methodology/data-processing-and-archiving/weighting).

A crucial qualification is that `brms::weights()` implements **likelihood weighting**: internally, each observation's log-posterior contribution is multiplied by its weight. It is therefore a weighted Bayesian pseudo-likelihood, not a full Bayesian treatment of the ESS's multistage sampling design. In particular, the posterior intervals should not automatically be interpreted as design-based survey uncertainty. The weights are used here to target the population represented by the ESS rather than to claim a complete design-based variance estimator. See the [brms documentation](https://paulbuerkner.com/brms/reference/brmsformula.html).

## Why use weights at all?

For these data, the answer is **yes** if the estimand is a population quantity rather than simply a description of the realized ESS respondents. The ESS explicitly recommends using weights for ESS analyses because respondents have unequal selection probabilities and because nonresponse, noncoverage, post-stratification, and differences in country population size affect representativeness. [ESS weighting guidance](https://www.europeansocialsurvey.org/sites/default/files/2023-06/ESS_weighting_data_1_1.pdf).

There are two separate questions:

- **Target population:** Who should the observations represent? For a pooled analysis of all ten countries as one population, `anweight` is appropriate.
- **Uncertainty:** How should sampling uncertainty be represented? `brms::weights()` does not reproduce the full ESS complex-sample variance calculation. This is why the analysis should be described as a weighted Bayesian pseudo-likelihood and why sensitivity checks against unweighted models are useful.

The weights should therefore not be regarded as an optional device for making the model fit better. They define the population to which the descriptive and substantive quantities refer.

# Setup

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
}

p_load(
  here, rio, tidyverse, brms, tidybayes, ggdist, posterior,
  bayesplot, tinytable, marginaleffects, modelsummary
)

options(
  dplyr.summarise.inform = FALSE,
  mc.cores = parallel::detectCores()
)

my_theme <- function(...) {
  theme(
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(color = "black", face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "grey70"),
    strip.text = element_text(color = "black"),
    legend.key.size = unit(0.3, "cm")
  )
}

set.seed(2026)

source(here("scripts", "_common.R"))

```

# Data and variable construction

```{r}
#| label: load-data

ess <- import(here("data", "processed", "ess_short_1_11.rds"), trust = TRUE)

countries <- c("BE", "CH", "DE", "ES", "FI", "FR", "NL", "NO", "PT", "SE")

# ESS Round 11 only. Round 10 was temporarily included in an earlier
# version of the analysis for diagnostic purposes, but is not part of this study.
ess11 <- ess |>
  filter(
    essround == 11,
    cntry %in% countries
  ) |>
  mutate(
    cntry = droplevels(factor(cntry)),

    socialist = case_when(
      vote_fct %in% c("Communist/Socialist", "Social democracy") ~ 1,
      vote_fct %in% c(
        "Agrarian", "Conservative/Christian democracy", "Green",
        "Liberal", "Radical right"
      ) ~ 0
    ),

    radicalright = case_when(
      vote_fct == "Radical right" ~ 1,
      vote_fct %in% c(
        "Agrarian", "Communist/Socialist",
        "Conservative/Christian democracy", "Green", "Liberal",
        "Social democracy"
      ) ~ 0
    ),

    # Education dummy, mirroring the class-based "worker" dummy used in the
    # parallel class-voting analysis. Reference category (0) is low education.
    eduhigh = case_when(
      edulowhigh == "high" ~ 1,
      edulowhigh == "low"  ~ 0
    )
  )

ess11 <- ess11 |>
  select(essround, cntry, age_grp_fct, edulowhigh, eduhigh,
        socialist, radicalright, married, gndr_fct, immigrant, pspwght, pweight) |>
  drop_na()

# Round 11 should contain the ESS analysis weight. If the processed file
# does not contain anweight, construct it from pspwght * pweight.
if ("anweight" %in% names(ess11)) {
  ess11 <- ess11 |>
    mutate(weight_an = anweight)
} else {
  ess11 <- ess11 |>
    mutate(weight_an = pspwght * pweight)
}

# Check the basic structure before estimation.
ess11 |>
  count(cntry) |>
  tt()

ess11 |>
  summarise(
    n = n(),
    n_missing_weight = sum(is.na(weight_an)),
    min_weight = min(weight_an, na.rm = TRUE),
    max_weight = max(weight_an, na.rm = TRUE),
    mean_weight = mean(weight_an, na.rm = TRUE)
  ) |>
  tt()
```

## Descriptive statistics of variables

```{r}
#| label: model-3-descriptives

vars <- c(
  "radicalright",
  "socialist",
  "eduhigh",
  "age_grp_fct",
  "gndr_fct",
  "married",
  "immigrant"
)

datasummary_skim(
  ess11[, vars],
  output = "tinytable",
  histogram = FALSE
)

```
## Construct the analysis sample

The main outcome is whether the respondent voted for a radical-right party among respondents for whom both vote choice and the education classification are observed.

```{r}
#| label: analysis-sample

voters <- ess11 |>
  filter(
    !is.na(radicalright),
    !is.na(eduhigh),
    !is.na(weight_an),
    weight_an > 0
  ) |>
  mutate(
    # Mean-one rescaling keeps the relative survey weights intact while
    # avoiding an arbitrary multiplication of the likelihood by a large constant.
    w = weight_an / mean(weight_an)
  )

voters |>
  summarise(
    n = n(),
    n_eff = sum(weight_an)^2 / sum(weight_an^2),
    mean_w = mean(w),
    min_w = min(w),
    max_w = max(w)
  ) |>
  tt()

voters |>
  summarise(
    n = n(),
    complete_m3 = sum(
      complete.cases(
        socialist,
        radicalright,
        eduhigh,
        age_grp_fct,
        gndr_fct,
        married,
        immigrant
      )
    )
  ) |>
  tt()

```

## A note on weight rescaling

The mean-one rescaling above is a modelling convention, not an ESS weighting convention. Multiplying all likelihood weights by a constant changes the amount of information supplied to a Bayesian pseudo-likelihood. Rescaling to mean one therefore makes the effective likelihood scale correspond roughly to the unweighted sample size while preserving the relative weights.

The Kish effective sample size is reported only as a diagnostic:

$$
N_{\mathrm{eff}} = \frac{(\sum_i w_i)^2}{\sum_i w_i^2}.
$$

It should **not** be used as the likelihood weight itself.

# Descriptive benchmark

Before fitting any model, calculate the two quantities directly from the weighted data. These provide a benchmark against which the posterior estimates can be checked.

```{r}
#| label: descriptive-benchmark

weighted_share <- function(x, w) {
  ok <- !is.na(x) & !is.na(w)
  weighted.mean(x[ok], w[ok])
}

# P(RR | high edu), P(RR | low edu), P(high edu), and P(high edu | RR)
benchmark <- voters |>
  summarise(
    p_rr_eduhigh = weighted.mean(radicalright[eduhigh == 1],
                                weight_an[eduhigh == 1], na.rm = TRUE),
    p_rr_edulow = weighted.mean(radicalright[eduhigh == 0],
                                   weight_an[eduhigh == 0], na.rm = TRUE),
    p_eduhigh = weighted.mean(eduhigh, weight_an, na.rm = TRUE),
    p_eduhigh_rr = weighted.mean(eduhigh[radicalright == 1],
                                weight_an[radicalright == 1], na.rm = TRUE)
  )

benchmark |>
  tt()
```

## Country-specific descriptive estimates

```{r}
#| label: country-benchmark

country_benchmark <- voters |>
  group_by(cntry) |>
  summarise(
    p_rr_eduhigh = weighted.mean(radicalright[eduhigh == 1],
                                weight_an[eduhigh == 1], na.rm = TRUE),
    p_rr_edulow = weighted.mean(radicalright[eduhigh == 0],
                                   weight_an[eduhigh == 0], na.rm = TRUE),
    p_eduhigh = weighted.mean(eduhigh, weight_an, na.rm = TRUE),
    p_eduhigh_rr = weighted.mean(eduhigh[radicalright == 1],
                                weight_an[radicalright == 1], na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(
    edu_gap = p_rr_eduhigh - p_rr_edulow
  )

country_benchmark |>
  tt()

```

# Model 0: pooled Bayesian model

The simplest useful model is:

$$
R_i \sim \operatorname{Bernoulli}(p_i)
$$

$$
\operatorname{logit}(p_i)=\alpha+\beta E_i.
$$

This one model estimates both conditional radical-right probabilities:

$$
P(RR\mid E=0)=\operatorname{logit}^{-1}(\alpha)
$$

and

$$
P(RR\mid E=1)=\operatorname{logit}^{-1}(\alpha+\beta).
$$

The contribution quantity is then derived by combining these probabilities with the weighted high-education prevalence:

$$
P(E\mid RR)=
\frac{P(RR\mid E)P(E)}
{P(RR\mid E)P(E)+P(RR\mid \neg E)P(\neg E)}.
$$

This is preferable to fitting a completely separate model for `P(high education | radical right)` because both quantities describe the same underlying contingency table.

## Priors

The priors below are deliberately weakly informative rather than flat. A normal prior of `normal(-2, 1)` on the intercept puts substantial prior mass on relatively low baseline probabilities of radical-right voting while still allowing much higher probabilities. The education coefficient has a symmetric prior because the direction of the association is not imposed by the prior.

```{r}
#| label: pooled-priors

pri_m0 <- c(
  prior(normal(-2, 1), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh")
)
```

## Fit the pooled weighted model

```{r}
#| label: pooled-model

m0 <- brm(
  radicalright | weights(w) ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_m0,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m0")
)

summary(m0)
```

## Derive loyalty and contribution from the posterior

We estimate `P(high education)` with a separate intercept-only weighted model. This is not a second model for contribution: it supplies the marginal high-education prevalence needed to transform the conditional radical-right probabilities into `P(high education | radical right)`.

```{r}
#| label: eduhigh-prevalence-model

m_eduhigh <- brm(
  eduhigh | weights(w) ~ 1,
  data = voters,
  family = bernoulli(),
  prior = prior(normal(0, 1), class = "Intercept"),
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m_eduhigh")
)

post_m0 <- as_draws_df(m0) |>
  as_tibble() |>
  select(.draw, alpha = b_Intercept, beta = b_eduhigh) |>
  inner_join(
    as_draws_df(m_eduhigh) |>
      as_tibble() |>
      select(.draw, gamma = b_Intercept),
    by = ".draw"
  ) |>
  mutate(
    p_rr_eduhigh = plogis(alpha + beta),
    p_rr_edulow = plogis(alpha),
    p_eduhigh = plogis(gamma),
    p_rr = p_rr_eduhigh * p_eduhigh +
      p_rr_edulow * (1 - p_eduhigh),
    p_eduhigh_rr = p_rr_eduhigh * p_eduhigh / p_rr,
    edu_gap = p_rr_eduhigh - p_rr_edulow
  )

post_m0 |>
  select(.draw, p_rr_eduhigh, p_rr_edulow, p_eduhigh,
         p_rr, p_eduhigh_rr, edu_gap) |>
  pivot_longer(-.draw, names_to = "quantity", values_to = "value") |>
  group_by(quantity) |>
  median_qi(value, .width = c(.8, .95)) |>
  tt()

```

## Posterior versus descriptive benchmark

```{r}
#| label: benchmark-comparison

posterior_summary_edu <- post_m0 |>
  summarise(
    p_rr_eduhigh = median(p_rr_eduhigh),
    p_rr_edulow = median(p_rr_edulow),
    p_eduhigh = median(p_eduhigh),
    p_eduhigh_rr = median(p_eduhigh_rr),
    edu_gap = median(edu_gap)
  )

bind_rows(
  benchmark |> mutate(source = "Weighted descriptive estimate"),
  posterior_summary_edu |> mutate(source = "Posterior median")
) |>
  relocate(source) |>
  tt()

```

# Model 1: country-varying intercepts

The next question is whether countries differ in their baseline propensity to vote for the radical right. The education effect is initially assumed to be common across countries:

$$
\operatorname{logit}(p_{ic})=\alpha_c+\beta E_{ic}.
$$

```{r}
#| label: model-1

pri_m1 <- c(
  prior(normal(-2, 1), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh"),
  prior(exponential(1), class = "sd")
)

m1 <- brm(
  radicalright | weights(w) ~ eduhigh + (1 | cntry),
  data = voters,
  family = bernoulli(),
  prior = pri_m1,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m1")
)

summary(m1)
```

# Model 2: country-varying education effects

The key substantive model allows the education association itself to vary across countries:

$$
\operatorname{logit}(p_{ic})=\alpha_c+\beta_c E_{ic}.
$$

The country-specific intercepts and slopes are given a joint multivariate normal distribution. This partial-pooling structure is preferable to estimating ten completely separate regressions because countries with smaller effective samples borrow information from the others.

```{r}
#| label: model-2

pri_m2 <- c(
  prior(normal(-2, 1), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh"),
  prior(exponential(1), class = "sd"),
  prior(lkj(2), class = "cor")
)

m2 <- brm(
  radicalright | weights(w) ~ eduhigh + (1 + eduhigh | cntry),
  data = voters,
  family = bernoulli(),
  prior = pri_m2,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m2")
)

summary(m2)

```

# Posterior predictions from Model 2

The main substantive quantities are easier to interpret on the probability scale than the log-odds coefficient. We therefore calculate, for each country:

- `P(RR | high education)`
- `P(RR | low education)`
- the absolute education gap.

```{r}
#| label: model-2-predictions

nd <- expand_grid(
  cntry = levels(voters$cntry),
  eduhigh = c(0, 1)
)

epred_m2 <- m2 |>
  epred_draws(
    newdata = nd,
    re_formula = NULL
  ) |>
  ungroup() |>
  select(cntry, eduhigh, .draw, .epred)

gap_m2 <- epred_m2 |>
  pivot_wider(
    names_from = eduhigh,
    values_from = .epred,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    edu_gap = eduhigh_1 - eduhigh_0
  )

# Posterior summary of the three central quantities.
gap_m2 |>
  pivot_longer(
    cols = c(eduhigh_1, eduhigh_0, edu_gap),
    names_to = "quantity",
    values_to = "value"
  ) |>
  group_by(cntry, quantity) |>
  median_qi(value, .width = c(.8, .95)) |>
  tt()

```

## Plot: loyalty by country

```{r}
#| label: plot-loyalty

ord <- epred_m2 |>
  filter(eduhigh == 1) |>
  group_by(cntry) |>
  summarise(m = median(.epred)) |>
  arrange(m) |>
  pull(cntry)

epred_m2 |>
  filter(eduhigh == 1) |>
  mutate(cntry = factor(cntry, levels = ord)) |>
  ggplot(aes(x = .epred, y = cntry)) +
  stat_halfeye(.width = c(.8, .95),
   fill = "grey80",
   color = "grey40") +
  scale_x_continuous(limits = c(0, .5)) +
  labs(
    x = "P(radical right | high education)",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals"
  )

ggsave(here("figs", "loyalty_eduhigh_rr_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")

```

## Plot: education gap

```{r}
#| label: plot-gap

gap_order <- gap_m2 |>
  group_by(cntry) |>
  summarise(m = median(edu_gap)) |>
  arrange(m) |>
  pull(cntry)

gap_m2 |>
  mutate(cntry = factor(cntry, levels = gap_order)) |>
  ggplot(aes(x = edu_gap, y = cntry)) +
  stat_halfeye(.width = c(.80, .95),
  point_interval = median_qi,
   fill  = "grey80",
   color = "grey40") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_x_continuous() +
  labs(
    x = "P(RR | high education) − P(RR | low education)",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals")

ggsave(here("figs", "loyalty_edugap_rr_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")

```

# Deriving contribution by country

The contribution quantity can be derived from the country-specific posterior predictions once we also estimate the country-specific prevalence of high-education respondents.

The identity is:

$$
P(E\mid RR,c)=
\frac{P(RR\mid E,c)P(E\mid c)}
{P(RR\mid E,c)P(E\mid c)+P(RR\mid \neg E,c)[1-P(E\mid c)]}.
$$

Rather than fitting a separate radical-right-only model for every country, estimate `P(high education | country)` with a hierarchical intercept model and combine the posterior predictions draw by draw.

```{r}
#| label: contribution-country
#| message: false
#| results: hide

m_eduhigh_country <- brm(
  eduhigh | weights(w) ~ 1 + (1 | cntry),
  data = voters,
  family = bernoulli(),
  prior = c(
    prior(normal(0, 1), class = "Intercept"),
    prior(exponential(1), class = "sd")
  ),
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m_eduhigh_country")
)

eduhigh_country <- m_eduhigh_country |>
  epred_draws(newdata = distinct(voters, cntry), re_formula = NULL) |>
  ungroup() |>
  select(cntry, .draw, p_eduhigh = .epred)

# Combine country-specific P(RR | high edu), P(RR | low edu), and P(high edu).
contribution_country <- gap_m2 |>
  rename(
    p_rr_edulow = eduhigh_0,
    p_rr_eduhigh = eduhigh_1
  ) |>
  inner_join(eduhigh_country, by = c("cntry", ".draw")) |>
  mutate(
    p_rr = p_rr_eduhigh * p_eduhigh +
      p_rr_edulow * (1 - p_eduhigh),
    p_eduhigh_rr = p_rr_eduhigh * p_eduhigh / p_rr
  )

contribution_country |>
  group_by(cntry) |>
  summarise(
    across(
      c(p_rr_eduhigh, p_rr_edulow, p_eduhigh, p_eduhigh_rr),
      list(median = median, lower = ~quantile(.x, .025), upper = ~quantile(.x, .975)),
      .names = "{.col}_{.fn}"
    )
  ) |>
tt()

```

## Plot of estimated contribution per country

The contribution estimate is the posterior probability that a voter has high education, conditional on voting for the radical right. We plot the full country-specific posterior distribution, with countries ordered by their posterior median.

```{r}
#| label: plot-contribution-country

contribution_order <- contribution_country |>
  group_by(cntry) |>
  summarise(m = median(p_eduhigh_rr)) |>
  arrange(m) |>
  pull(cntry)

contribution_country |>
  mutate(cntry = factor(cntry, levels = contribution_order)) |>
  ggplot(aes(x = p_eduhigh_rr, y = cntry)) +
  stat_halfeye(
    point_interval = median_qi,
    .width = c(.80, .95),
    fill = "grey80",
    color = "grey40"
  ) +
  scale_x_continuous() +
  labs(
    x = "P(high education | radical right)",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals"
  )

ggsave(
  here("figs", "contribution_eduhigh_rr_countries.png"),
  width = 10,
  height = 7,
  dpi = 300,
  bg = "white"
)
```

# Posterior predictive checks

Posterior predictive checks are used to ask whether the models can reproduce important features of the observed data. Because the pooled model ignores country heterogeneity, country-specific checks are particularly useful for deciding whether Model 1 and Model 2 are warranted.

```{r}
#| label: pp-check-country

set.seed(2026)

# For large weighted samples, use a moderate number of posterior draws.
yrep <- posterior_predict(m0, ndraws = 300)

rep_means <- map_dfr(levels(voters$cntry), function(cc) {
  idx <- voters$cntry == cc
  tibble(
    cntry = cc,
    .draw = seq_len(nrow(yrep)),
    rep_mean = rowMeans(yrep[, idx, drop = FALSE])
  )
})

obs_means <- voters |>
  group_by(cntry) |>
  summarise(
    obs_mean = weighted.mean(radicalright, weight_an, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(rep_means, aes(rep_mean)) +
  geom_histogram(bins = 30) +
  geom_vline(
    data = obs_means,
    aes(xintercept = obs_mean),
    linewidth = 0.8
  ) +
  facet_wrap(~ cntry, scales = "free_x") +
  labs(
    x = "Mean radical-right vote in replicated data",
    y = NULL,
    title = "Pooled model: country-specific posterior predictive check",
    subtitle = "Vertical lines show weighted observed country means"
  )

```

# Prior predictive check

Before interpreting the fitted models, it is useful to check whether the priors imply plausible radical-right voting probabilities. This is especially important because a logistic intercept prior translates nonlinearly into the probability scale.

```{r}
#| label: prior-predictive
#| message: false
#| results: hide

m0_prior <- brm(
  radicalright ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_m0,
  sample_prior = "only",
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 1000,
  warmup = 500,
  seed = 2026,
  file = here("models", "ess11_edu_m0_prior")
)

prior_epred <- m0_prior |>
  epred_draws(
    newdata = tibble(eduhigh = c(0, 1)),
    ndraws = 500
  ) |>
  mutate(eduhigh = ifelse(eduhigh == 0, "Low education", "High education"))

ggplot(prior_epred, aes(x = .epred)) +
  stat_halfeye() +
  facet_wrap(~ eduhigh) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    x = "Prior predictive P(radical right)",
    y = NULL,
    title = "Prior predictive distribution"
  )


```

# Sensitivity to weighting

The main estimates use ESS weights because the substantive estimands concern population quantities. Nevertheless, it is useful to fit the simplest model without weights as a **sensitivity analysis**.

The comparison should not be interpreted as choosing whichever model gives the preferred result. Rather, it tells us whether the substantive conclusion depends materially on the survey weighting.

```{r}
#| label: weighting-sensitivity
#| message: false
#| results: hide

m0_unweighted <- brm(
  radicalright ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_m0,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m0_unweighted")
)

# Compare the central education-group probabilities.
weighted_draws <- as_draws_df(m0) |>
  as_tibble() |>
  transmute(
    model = "Weighted",
    p_rr_edulow = plogis(b_Intercept),
    p_rr_eduhigh = plogis(b_Intercept + b_eduhigh)
  )

unweighted_draws <- as_draws_df(m0_unweighted) |>
  as_tibble() |>
  transmute(
    model = "Unweighted",
    p_rr_edulow = plogis(b_Intercept),
    p_rr_eduhigh = plogis(b_Intercept + b_eduhigh)
  )

bind_rows(weighted_draws, unweighted_draws) |>
  pivot_longer(
    cols = starts_with("p_rr"),
    names_to = "quantity",
    values_to = "value"
  ) |>
  group_by(model, quantity) |>
  median_qi(value, .width = c(.8, .95)) |>
  tt()

```

# Optional sensitivity: country-balanced weighting

The default pooled ESS analysis weights countries in proportion to their population size. That is appropriate if the estimand is a population-level quantity for the ten countries taken together.

A different estimand would give each country equal population-level importance. This is not the ESS default pooled estimand. If the substantive question is explicitly about the **average country** rather than the population of all ten countries, country-normalised weights can be used as a sensitivity analysis.

```{r}
#| label: country-balanced-weights

voters_country_balanced <- voters |>
  group_by(cntry) |>
  mutate(
    w_country_balanced = weight_an / mean(weight_an)
  ) |>
  ungroup()

# Check that every country now contributes the same total weight.
voters_country_balanced |>
  group_by(cntry) |>
  summarise(total_weight = sum(w_country_balanced)) |>
  tt()

```

# Adding covariates: next stage

Only after the education association and country heterogeneity have been established should additional individual-level variables be introduced.

The proposed sequence is:

### Model 3: compositional adjustment

```r
radicalright | weights(w) ~ eduhigh + cohort_fct + gndr_fct + married + immigrant + (1 + eduhigh | cntry)
```

This asks how the education association changes after accounting for differences in the composition of the high- and low-education groups.

The factor reference categories are set explicitly so that the Model 3 coefficients have transparent interpretations. The reference groups are 18–29-year-olds, men, the reference marital-status category “at one point”, non-immigrants, and respondents with low education (below upper secondary). The `get_prior()` call is then used to verify the exact coefficient names before specifying the priors.

```{r}
#| label: model-3-references-priors
#| message: false
#| results: hide

voters_m3 <- voters |>
  mutate(
    age_grp_fct = relevel(factor(age_grp_fct), ref = "18-29"),
    gndr_fct    = relevel(factor(gndr_fct), ref = "men"),
    married     = relevel(factor(married), ref = "at one point")
  )

get_prior(
  radicalright | weights(w) ~
    eduhigh +
    age_grp_fct +
    gndr_fct +
    married +
    immigrant +
    (1 + eduhigh | cntry),
  data = voters_m3,
  family = bernoulli()
)

pri_m3 <- c(
  # Baseline probability for 18–29-year-old, male, non-immigrant,
  # low-education respondent in the reference marital-status category
  prior(normal(-2.3, 0.7), class = "Intercept"),

  # High vs. low education
  prior(normal(0, 0.5), class = "b", coef = "eduhigh"),

  # Age groups
  prior(normal(0, 0.5), class = "b", coef = "age_grp_fct30M65"),
  prior(normal(0, 0.5), class = "b", coef = "age_grp_fctabove65"),

  # Gender, immigration, and marital status
  prior(normal(0, 0.5), class = "b", coef = "gndr_fctwomen"),
  prior(normal(0, 0.5), class = "b", coef = "immigrant"),
  prior(normal(0, 0.5), class = "b", coef = "marriednever"),

  # Country-level variation
  prior(exponential(2), class = "sd"),
  prior(lkj(2), class = "cor")
)

# prior predictive check
pp_check(
  brm(
    radicalright ~
      eduhigh +
      age_grp_fct +
      gndr_fct +
      married +
      immigrant +
      (1 + eduhigh | cntry),
    data = voters_m3,
    family = bernoulli(),
    prior = pri_m3,
    sample_prior = "only",
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  file = here("models", "ess11_edu_pp_check")
  )
)

```

After the prior predictive check has been inspected and found reasonable, Model 3 is fitted with the same country-varying intercept and education slope structure as Model 2, while adding age group, gender, marital status, and immigrant status. The primary estimand remains the education difference, so the additional covariates are treated as a compositional adjustment rather than as a causal identification strategy.

```{r}
#| label: model-3-fit
#| message: false
#| results: hide

m3 <- brm(
  radicalright | weights(w) ~
    eduhigh +
    age_grp_fct +
    gndr_fct +
    married +
    immigrant +
    (1 + eduhigh | cntry),
  data = voters_m3,
  family = bernoulli(),
  prior = pri_m3,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m3")
)

summary(m3)
plot(m3)
pp_check(m3, type = "bars")

```

## Plot coefficients

The coefficient plot provides a compact view of the adjusted associations on the log-odds scale. It is useful for seeing the direction and uncertainty of the individual covariate coefficients, but it is not the main substantive presentation of Model 3. The latter is given by standardized predicted probabilities and their education difference below.

```{r}
#| label: model-3-coefficients

draws <- as_draws_df(m3)

draws <- draws |>
  rename(
    "High education" = b_eduhigh,
    "Age 30-65" = b_age_grp_fct30M65,
    "Age 65+"   = b_age_grp_fctabove65,
    Woman = b_gndr_fctwomen,
    Immigrant = "b_immigrant" )

mcmc_areas(
  draws,
  pars = c("High education", "Age 30-65", "Age 65+", "Woman", "Immigrant"),
  prob = 0.90)+
  geom_vline(xintercept = 0,
             linetype = "dotted")

```

## Average difference in predicted probability

The main substantive quantity is calculated on the probability scale. We duplicate the observed sample, set `eduhigh` successively to 0 and 1, and retain the observed distribution of the adjustment variables. Predictions exclude the country-specific random effects here (`re_formula = NA`) so that this first comparison represents the population-level relationship standardized over the observed covariate distribution.

```{r}
#| label: model-3-standardized-predictions

newdata_m3 <- bind_rows(
  voters_m3 |> mutate(eduhigh = 0),
  voters_m3 |> mutate(eduhigh = 1)
) |>
  mutate(.row_weight = w)

posterior_m3 <- m3 |>
  add_epred_draws(
    newdata = newdata_m3,
    re_formula = NA
  )

# Calculate the weighted standardized probability for each education group
# within each posterior draw.
std_m3 <- posterior_m3 |>
  group_by(.draw, eduhigh) |>
  summarise(
    p = weighted.mean(.epred, w),
    .groups = "drop"
  )

# Convert the two standardized probabilities into their posterior difference.
std_diff_m3 <- std_m3 |>
  select(.draw, eduhigh, p) |>
  tidyr::pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

ggplot(std_diff_m3, aes(x = diff)) +
  stat_halfeye(.width = c(.8, .95),
   fill  = "grey80",
   color = "grey40") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Difference in predicted probability \n Posterior distribution",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals")

```

## Country-specific adjusted predictions

The pooled standardized prediction above describes the average education difference. We next retain the country-specific random effects to examine whether the adjusted difference varies across countries. Within each country, high- and low-education respondents are compared using the same observed distribution of age, gender, marital status, and immigrant status.

```{r}
#| label: model-3-country-predictions

newdata_m3_country <- bind_rows(
  voters_m3 |> mutate(eduhigh = 0),
  voters_m3 |> mutate(eduhigh = 1)
)

posterior_m3_country <- m3 |>
  add_epred_draws(
    newdata = newdata_m3_country,
    re_formula = NULL
  )

std_m3_country <- posterior_m3_country |>
  group_by(.draw, cntry, eduhigh) |>
  summarise(
    p = weighted.mean(.epred, w),
    .groups = "drop"
  )

diff_m3_country <- std_m3_country |>
  pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

diff_m3_country <- std_m3_country |>
  tidyr::pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

country_diff <- diff_m3_country |>
group_by(cntry) |>
summarise(diff_med = median(diff))

std_m3_country |>
left_join(country_diff, by = "cntry") |>
ggplot(
  aes(
    x = p,
    y = forcats::fct_reorder(cntry, diff_med),
    colour = factor(eduhigh)
  )
) +
  stat_pointinterval(
    point_interval = median_qi,
    .width = c(0.8, 0.95),
    position = position_dodge(width = 0.5)
  ) +
  scale_x_continuous(limits = c(0, 0.4)) +
  scale_colour_discrete(
    name = NULL,
    labels = c("Low education", "High education")
  ) +
  labs(
    x = "Predicted probability of radical-right support",
    y = NULL
  )

ggsave(here("figs", "pp_diff_edu_rr_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")

```


## Interpreting the Model 3 predictions

The pooled standardized probabilities provide the primary estimate of the education difference after compositional adjustment. The country-specific predictions show whether that adjusted association remains heterogeneous across countries.

The most informative comparison with Model 2 is the posterior distribution of the country-level standard deviation of the education effect, `sd_cntry__eduhigh`. A reduction relative to Model 2 would indicate that some of the previously estimated cross-national heterogeneity is associated with differences in the observed composition of the high- and low-education groups. Persistent variation would indicate that country-level heterogeneity remains after this adjustment.

These comparisons are descriptive rather than causal. The Model 3 covariates are not assumed to satisfy the conditions required for causal confounding adjustment.

```{r}
#| label: modelcomparison

posterior_summary(
  m3,
  variable = c(
    "b_eduhigh",
    "sd_cntry__eduhigh"
  )
)

posterior_summary(
  m2,
  variable = c(
    "b_eduhigh",
    "sd_cntry__eduhigh"
  )
)

```

## Calculate the posterior distribution of the standardized high–low education difference in predicted probability.
```{r}
#| label: pp_eduhigh_edulow

std_diff_m3 <- std_m3 |>
  tidyr::pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

std_diff_m3 |>
  summarise(
    median = median(diff),
    lower  = quantile(diff, 0.025),
    upper  = quantile(diff, 0.975)
  )

# Plot the posterior distribution of the standardized
# high-education minus low-education probability difference.

ggplot(std_diff_m3, aes(x = diff)) +
  stat_halfeye(
    .width = c(0.8, 0.95)
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed"
  ) +
  scale_x_continuous(
    labels = scales::label_percent()
  ) +
  labs(
    x = "High − low education difference in predicted probability",
    y = NULL,
    caption = "Model 3; predictions standardised to the observed covariate distribution"
  )

```

# Socialist

The section repeats the above analysis for socialist parties, defined as **Social democracy** plus **Communist/Socialist** parties. The outcome is `socialist`, which was already constructed in the data-preparation section.

The main difference from the radical-right analysis is the prior on the intercept. Socialist voting is substantially more common than radical-right voting in these data, so the radical-right intercept prior would be unnecessarily restrictive on the probability scale. The priors below therefore shift the intercept towards a higher baseline probability while retaining the same weakly informative structure for the education association and the country-level heterogeneity.

## Descriptive benchmark

Before fitting the Bayesian models, calculate the weighted socialist vote among high- and low-education respondents. These estimates provide a useful empirical benchmark for the prior predictive checks and for the subsequent posterior predictions.

```{r}
#| label: socialist-descriptive-benchmark

socialist_benchmark <- voters |>
  summarise(
    p_soc_eduhigh = weighted.mean(
      socialist[eduhigh == 1],
      weight_an[eduhigh == 1],
      na.rm = TRUE
    ),
    p_soc_edulow = weighted.mean(
      socialist[eduhigh == 0],
      weight_an[eduhigh == 0],
      na.rm = TRUE
    ),
    p_eduhigh = weighted.mean(
      eduhigh,
      weight_an,
      na.rm = TRUE
    ),
    p_eduhigh_soc = weighted.mean(
      eduhigh[socialist == 1],
      weight_an[socialist == 1],
      na.rm = TRUE
    )
  ) |>
  mutate(
    edu_gap = p_soc_eduhigh - p_soc_edulow
  )

socialist_benchmark |> tt()
```

## Country-specific descriptive estimates

```{r}
#| label: socialist-country-benchmark

socialist_country_benchmark <- voters |>
  group_by(cntry) |>
  summarise(
    p_soc_eduhigh = weighted.mean(
      socialist[eduhigh == 1],
      weight_an[eduhigh == 1],
      na.rm = TRUE
    ),
    p_soc_edulow = weighted.mean(
      socialist[eduhigh == 0],
      weight_an[eduhigh == 0],
      na.rm = TRUE
    ),
    p_eduhigh = weighted.mean(
      eduhigh,
      weight_an,
      na.rm = TRUE
    ),
    p_eduhigh_soc = weighted.mean(
      eduhigh[socialist == 1],
      weight_an[socialist == 1],
      na.rm = TRUE
    ),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(
    edu_gap = p_soc_eduhigh - p_soc_edulow
  )

socialist_country_benchmark |> tt()
```

# Model 0: pooled Bayesian model for socialist voting

The pooled model is:

$$
S_i \sim \operatorname{Bernoulli}(p_i)
$$

$$
\operatorname{logit}(p_i)=\alpha+\beta E_i.
$$

This gives:

$$
P(S\mid E=0)=\operatorname{logit}^{-1}(\alpha)
$$

and

$$
P(S\mid E=1)=\operatorname{logit}^{-1}(\alpha+\beta).
$$

The contribution quantity is derived from these probabilities and the posterior distribution of the high-education prevalence.

## Priors

The socialist intercept prior is shifted upwards relative to the radical-right prior. `normal(-1.1, 0.6)` has a median probability of about 25% on the probability scale and still allows substantial uncertainty around that value. This is intended to reflect the much higher prevalence of socialist voting without making the prior depend tightly on the observed outcome.

The education coefficient remains centred at zero because the prior should not determine the direction of the education association.

```{r}
#| label: socialist-pooled-priors

pri_soc_m0 <- c(
  prior(normal(-1.1, 0.6), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh")
)
```

## Fit the pooled weighted model

```{r}
#| label: socialist-pooled-model

m0_soc <- brm(
  socialist | weights(w) ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_soc_m0,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m0_soc")
)

summary(m0_soc)
```

## Derive socialist loyalty and contribution

The education-prevalence model from the radical-right analysis can be reused because `P(high education)` is a property of the sample, not of the party outcome. The socialist posterior predictions are then combined with the same posterior draws of education prevalence.

```{r}
#| label: socialist-pooled-predictions

post_soc_m0 <- as_draws_df(m0_soc) |>
  as_tibble() |>
  select(.draw, alpha = b_Intercept, beta = b_eduhigh) |>
  inner_join(
    as_draws_df(m_eduhigh) |>
      as_tibble() |>
      select(.draw, gamma = b_Intercept),
    by = ".draw"
  ) |>
  mutate(
    p_soc_eduhigh = plogis(alpha + beta),
    p_soc_edulow = plogis(alpha),
    p_eduhigh = plogis(gamma),
    p_soc = p_soc_eduhigh * p_eduhigh +
      p_soc_edulow * (1 - p_eduhigh),
    p_eduhigh_soc = p_soc_eduhigh * p_eduhigh / p_soc,
    edu_gap = p_soc_eduhigh - p_soc_edulow
  )

post_soc_m0 |>
  select(
    .draw,
    p_soc_eduhigh,
    p_soc_edulow,
    p_eduhigh,
    p_soc,
    p_eduhigh_soc,
    edu_gap
  ) |>
  pivot_longer(
    - .draw,
    names_to = "quantity",
    values_to = "value"
  ) |>
  group_by(quantity) |>
  median_qi(value, .width = c(.80, .95))

```

## Posterior versus descriptive benchmark

```{r}
#| label: socialist-benchmark-comparison

socialist_posterior_summary <- post_soc_m0 |>
  summarise(
    p_soc_eduhigh = median(p_soc_eduhigh),
    p_soc_edulow = median(p_soc_edulow),
    p_eduhigh = median(p_eduhigh),
    p_eduhigh_soc = median(p_eduhigh_soc),
    edu_gap = median(edu_gap)
  )

bind_rows(
  socialist_benchmark |>
    mutate(source = "Weighted descriptive estimate"),
  socialist_posterior_summary |>
    mutate(source = "Posterior median")
) |>
  relocate(source) |>
  tt()
```

# Prior predictive check for socialist voting

The prior predictive check is especially important here because the intercept prior has been shifted to reflect the substantially higher prevalence of socialist voting.

```{r}
#| label: socialist-prior-predictive

m0_soc_prior <- brm(
  socialist ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_soc_m0,
  sample_prior = "only",
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 1000,
  warmup = 500,
  seed = 2026,
  file = here("models", "ess11_edu_m0_soc_prior")
)

prior_soc_epred <- m0_soc_prior |>
  epred_draws(
    newdata = tibble(eduhigh = c(0, 1)),
    ndraws = 500
  ) |>
  mutate(
    eduhigh = ifelse(eduhigh == 0, "Low education", "High education")
  )

ggplot(prior_soc_epred, aes(x = .epred)) +
  stat_halfeye() +
  facet_wrap(~ eduhigh) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    x = "Prior predictive P(socialist vote)",
    y = NULL,
    title = "Prior predictive distribution for socialist voting"
  )

```

# Model 1: country-varying intercepts

The first extension allows countries to differ in their baseline level of socialist voting:

$$
\operatorname{logit}(p_{ic})=\alpha_c+\beta E_{ic}.
$$

The intercept prior is again shifted relative to the radical-right model.

```{r}
#| label: socialist-model-1

pri_soc_m1 <- c(
  prior(normal(-1.1, 0.6), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh"),
  prior(exponential(1), class = "sd")
)

m1_soc <- brm(
  socialist | weights(w) ~ eduhigh + (1 | cntry),
  data = voters,
  family = bernoulli(),
  prior = pri_soc_m1,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m1_soc")
)

summary(m1_soc)
```

# Model 2: country-varying education effects

The main hierarchical model allows both the baseline level of socialist voting and the education association to vary across countries:

$$
\operatorname{logit}(p_{ic})=\alpha_c+\beta_c E_{ic}.
$$

```{r}
#| label: socialist-model-2

pri_soc_m2 <- c(
  prior(normal(-1.1, 0.6), class = "Intercept"),
  prior(normal(0, 1), class = "b", coef = "eduhigh"),
  prior(exponential(1), class = "sd"),
  prior(lkj(2), class = "cor")
)

m2_soc <- brm(
  socialist | weights(w) ~ eduhigh + (1 + eduhigh | cntry),
  data = voters,
  family = bernoulli(),
  prior = pri_soc_m2,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m2_soc")
)

summary(m2_soc)
```

# Posterior predictions from Model 2

```{r}
#| label: socialist-model-2-predictions

nd_soc <- expand_grid(
  cntry = levels(voters$cntry),
  eduhigh = c(0, 1)
)

epred_soc_m2 <- m2_soc |>
  epred_draws(
    newdata = nd_soc,
    re_formula = NULL
  ) |>
  ungroup() |>
  select(cntry, eduhigh, .draw, .epred)

gap_soc_m2 <- epred_soc_m2 |>
  pivot_wider(
    names_from = eduhigh,
    values_from = .epred,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    edu_gap = eduhigh_1 - eduhigh_0
  )

gap_soc_m2 |>
  pivot_longer(
    cols = c(eduhigh_1, eduhigh_0, edu_gap),
    names_to = "quantity",
    values_to = "value"
  ) |>
  group_by(cntry, quantity) |>
  median_qi(value, .width = c(.8, .95)) |>
  tt()

```

## Plot: socialist loyalty by country

```{r}
#| label: socialist-plot-loyalty

ord_soc <- epred_soc_m2 |>
  filter(eduhigh == 1) |>
  group_by(cntry) |>
  summarise(m = median(.epred)) |>
  arrange(m) |>
  pull(cntry)

epred_soc_m2 |>
  filter(eduhigh == 1) |>
  mutate(cntry = factor(cntry, levels = ord_soc)) |>
  ggplot(aes(x = .epred, y = cntry)) +
  stat_halfeye(.width = c(.8, .95),
    color = "grey40",
    fill  = "grey80") +
  scale_x_continuous(limits = c(0, 1)) +
  labs(
    x = "P(socialist vote | high education)",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals"
  )

ggsave(here("figs", "loyalty_eduhigh_soc_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")
```

## Plot: high/low education gap

```{r}
#| label: socialist-plot-gap

gap_soc_order <- gap_soc_m2 |>
  group_by(cntry) |>
  summarise(m = median(edu_gap)) |>
  arrange(m) |>
  pull(cntry)

gap_soc_m2 |>
  mutate(cntry = factor(cntry, levels = gap_soc_order)) |>
  ggplot(aes(x = edu_gap, y = cntry)) +
  stat_halfeye(.width = c(.8, .95), point_interval = median_qi,
    color = "grey40",
    fill  = "grey80"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "P(socialist | high education) − P(socialist | low education)",
    y = NULL )

ggsave(here("figs", "loyalty_edugap_soc_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")
```

# Deriving socialist contribution by country

The contribution quantity is:

$$
P(E\mid S,c)=
\frac{P(S\mid E,c)P(E\mid c)}
{P(S\mid E,c)P(E\mid c)+P(S\mid \neg E,c)[1-P(E\mid c)]}.
$$

The country-specific education-prevalence model from the radical-right analysis can be reused because it does not depend on the party outcome.

```{r}
#| label: socialist-contribution-country

eduhigh_country_soc <- eduhigh_country

contribution_soc_country <- gap_soc_m2 |>
  rename(
    p_soc_edulow = eduhigh_0,
    p_soc_eduhigh = eduhigh_1
  ) |>
  inner_join(
    eduhigh_country_soc,
    by = c("cntry", ".draw")
  ) |>
  mutate(
    p_soc = p_soc_eduhigh * p_eduhigh +
      p_soc_edulow * (1 - p_eduhigh),
    p_eduhigh_soc = p_soc_eduhigh * p_eduhigh / p_soc
  )

contribution_soc_country |>
  group_by(cntry) |>
  summarise(
    across(
      c(
        p_soc_eduhigh,
        p_soc_edulow,
        p_eduhigh,
        p_eduhigh_soc
      ),
      list(
        median = median,
        lower = ~ quantile(.x, 0.025),
        upper = ~ quantile(.x, 0.975)
      ),
      .names = "{.col}_{.fn}"
    )
  ) |>
  tt()
```

## Plot of estimated contribution per country

The contribution estimate is the posterior probability that a voter has high education, conditional on voting for a socialist party. Countries are ordered by their posterior median.

```{r}
#| label: socialist-plot-contribution-country

contribution_soc_order <- contribution_soc_country |>
  group_by(cntry) |>
  summarise(m = median(p_eduhigh_soc)) |>
  arrange(m) |>
  pull(cntry)

contribution_soc_country |>
  mutate(cntry = factor(cntry, levels = contribution_soc_order)) |>
  ggplot(aes(x = p_eduhigh_soc, y = cntry)) +
  stat_halfeye(
    point_interval = median_qi,
    .width = c(.80, .95),
    fill  = "grey80",
    color = "grey40"
  ) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1)
  ) +
  labs(
    x = "P(high education | socialist)",
    y = NULL,
    caption = "Posterior median with 80% and 95% credible intervals"
  )

ggsave(
  here("figs", "contribution_eduhigh_soc_countries.png"),
  width = 10,
  height = 7,
  dpi = 300,
  bg = "white"
)
```

# Model 3: compositional adjustment

Model 3 repeats the covariate adjustment used for radical-right voting. The only substantive change is the outcome and the intercept prior.

The intercept prior is shifted from `normal(-2.3, 0.7)` to `normal(-1.4, 0.6)`. The tighter scale reflects the fact that the baseline probability for the reference group should still be plausible on the higher socialist-vote scale, while allowing substantial uncertainty.

```{r}
#| label: socialist-model-3-references-priors

voters_soc_m3 <- voters |>
  mutate(
    age_grp_fct = relevel(factor(age_grp_fct), ref = "18-29"),
    gndr_fct    = relevel(factor(gndr_fct), ref = "men"),
    married     = relevel(factor(married), ref = "at one point")
  )

get_prior(
  socialist | weights(w) ~
    eduhigh +
    age_grp_fct +
    gndr_fct +
    married +
    immigrant +
    (1 + eduhigh | cntry),
  data = voters_soc_m3,
  family = bernoulli()
)

pri_soc_m3 <- c(
  # Higher baseline socialist-vote probability than in the radical-right model
  prior(normal(-.6, 0.4), class = "Intercept"),

  # High vs. low education
  prior(normal(0, 0.5), class = "b", coef = "eduhigh"),

  # Age groups
  prior(normal(0, 0.5), class = "b", coef = "age_grp_fct30M65"),
  prior(normal(0, 0.5), class = "b", coef = "age_grp_fctabove65"),

  # Gender, immigration, and marital status
  prior(normal(0, 0.5), class = "b", coef = "gndr_fctwomen"),
  prior(normal(0, 0.5), class = "b", coef = "immigrant"),
  prior(normal(0, 0.5), class = "b", coef = "marriednever"),

  # Country-level variation
  prior(exponential(2), class = "sd"),
  prior(lkj(2), class = "cor")
)

```

## Prior predictive check for Model 3

```{r}
#| label: socialist-model-3-prior-predictive

m3_soc_prior <- brm(
  socialist ~
    eduhigh +
    age_grp_fct +
    gndr_fct +
    married +
    immigrant +
    (1 + eduhigh | cntry),
  data = voters_soc_m3,
  family = bernoulli(),
  prior = pri_soc_m3,
  sample_prior = "only",
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  file = here("models", "ess11_edu_m3_soc_prior")
)
pp_check(m3_soc_prior, type = "bars")

prior_epred_soc <- m3_soc_prior |>
  epred_draws(
    newdata = voters,
    re_formula = NULL,
    ndraws = 500
  )

ggplot(prior_epred_soc, aes(x = .epred)) +
  geom_histogram(bins = 50) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1)
  ) +
  labs(
    x = "Prior predicted probability of voting socialist",
    y = "Count"
  )

prior_epred_soc |>
  group_by(.draw) |>
  summarise(
    mean_p = mean(.epred),
    .groups = "drop"
  ) |>
  ggplot(aes(x = mean_p)) +
  geom_histogram(bins = 40) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1)
  ) +
  labs(
    x = "Mean prior predicted probability of voting socialist",
    y = "Number of prior draws"
  )



```

## Fit Model 3

```{r}
#| label: socialist-model-3-fit

m3_soc <- brm(
  socialist | weights(w) ~
    eduhigh +
    age_grp_fct +
    gndr_fct +
    married +
    immigrant +
    (1 + eduhigh | cntry),
  data = voters_soc_m3,
  family = bernoulli(),
  prior = pri_soc_m3,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95),
  file = here("models", "ess11_edu_m3_soc")
)

summary(m3_soc)
#plot(m3_soc)
pp_check(m3_soc, type = "bars")

```

## Standardized high/low education predictions

```{r}
#| label: socialist-model-3-standardized-predictions

newdata_soc_m3 <- bind_rows(
  voters_soc_m3 |> mutate(eduhigh = 0),
  voters_soc_m3 |> mutate(eduhigh = 1)
) |>
  mutate(.row_weight = w)

posterior_soc_m3 <- m3_soc |>
  add_epred_draws(
    newdata = newdata_soc_m3,
    re_formula = NA
  )

std_soc_m3 <- posterior_soc_m3 |>
  group_by(.draw, eduhigh) |>
  summarise(
    p = weighted.mean(.epred, w),
    .groups = "drop"
  )

std_diff_soc_m3 <- std_soc_m3 |>
  pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

std_diff_soc_m3 |>
  summarise(
    median = median(diff),
    lower = quantile(diff, 0.025),
    upper = quantile(diff, 0.975)
  )

ggplot(std_diff_soc_m3, aes(x = diff)) +
  stat_halfeye(.width = c(.80, .95),
    color = "grey40",
    fill  = "grey80") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_x_continuous() +
  labs(
    x = "High − low education difference in predicted probability",
    y = NULL  )

ggsave(
  here("figs", "dif_eduhigh_edulow_soc.png"),
  width = 10,
  height = 7,
  dpi = 300,
  bg = "white"
)

```

## Country-specific adjusted predictions

```{r}
#| label: socialist-model-3-country-predictions

newdata_m3_country <- bind_rows(
  voters_soc_m3 |> mutate(eduhigh = 0),
  voters_soc_m3 |> mutate(eduhigh = 1)
)

posterior_m3_country <- m3_soc |>
  add_epred_draws(
    newdata = newdata_m3_country,
    re_formula = NULL
  )

std_m3_country <- posterior_m3_country |>
  group_by(.draw, cntry, eduhigh) |>
  summarise(
    p = weighted.mean(.epred, w),
    .groups = "drop"
  )

diff_m3_country <- std_m3_country |>
  tidyr::pivot_wider(
    names_from = eduhigh,
    values_from = p,
    names_prefix = "eduhigh_"
  ) |>
  mutate(
    diff = eduhigh_1 - eduhigh_0
  )

country_diff <- diff_m3_country |>
group_by(cntry) |>
summarise(diff_med = median(diff))

std_m3_country |>
left_join(country_diff, by = "cntry") |>
ggplot(
  aes(
    x = p,
    y = forcats::fct_reorder(cntry, diff_med),
    colour = factor(eduhigh)
  )
) +
  stat_pointinterval(
    point_interval = median_qi,
    .width = c(0.8, 0.95),
    position = position_dodge(width = 0.5)
  ) +
  scale_x_continuous(limits = c(0, 0.8)) +
  scale_colour_discrete(
    name = NULL,
    labels = c("Low education", "High education")
  ) +
  labs(
    x = "Predicted probability of radical-right support",
    y = NULL
  )

ggsave(here("figs", "pp_diff_edu_soc_countries.png"),
      width = 10, height = 7, dpi = 300,
      bg = "white")

```

## Model comparison

As in the radical-right analysis, compare the education coefficient and the cross-country standard deviation of the education effect across Models 2 and 3.

```{r}
#| label: socialist-model-comparison

posterior_summary(
  m3_soc,
  variable = c(
    "b_eduhigh",
    "sd_cntry__eduhigh"
  )
)

posterior_summary(
  m2_soc,
  variable = c(
    "b_eduhigh",
    "sd_cntry__eduhigh"
  )
)
```

## Standardized high–low education difference

```{r}
#| label: socialist-pp-eduhigh-edulow

std_diff_soc_m3 |>
  summarise(
    median = median(diff),
    lower  = quantile(diff, 0.025),
    upper  = quantile(diff, 0.975)
  )

ggplot(std_diff_soc_m3, aes(x = diff)) +
  stat_halfeye(.width = c(0.80, 0.95),
    color = "grey40",
    fill  = "grey80") +
  geom_vline(
    xintercept = 0,
    linetype = "dashed"
  ) +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(
    x = "High − low education difference in predicted probability",
    y = NULL,
    caption = "Model 3; predictions standardised to the observed covariate distribution"
  )

```

# Weighting sensitivity for socialist voting

The same weighting sensitivity analysis can be repeated for socialist voting.

```{r}
#| label: socialist-weighting-sensitivity

m0_soc_unweighted <- brm(
  socialist ~ eduhigh,
  data = voters,
  family = bernoulli(),
  prior = pri_soc_m0,
  backend = "cmdstanr",
  threads = threading(4),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 2026,
  control = list(adapt_delta = 0.95)
)

weighted_soc_draws <- as_draws_df(m0_soc) |>
  as_tibble() |>
  transmute(
    model = "Weighted",
    p_soc_edulow = plogis(b_Intercept),
    p_soc_eduhigh = plogis(b_Intercept + b_eduhigh)
  )

unweighted_soc_draws <- as_draws_df(m0_soc_unweighted) |>
  as_tibble() |>
  transmute(
    model = "Unweighted",
    p_soc_edulow = plogis(b_Intercept),
    p_soc_eduhigh = plogis(b_Intercept + b_eduhigh)
  )

bind_rows(weighted_soc_draws, unweighted_soc_draws) |>
  pivot_longer(
    cols = starts_with("p_soc"),
    names_to = "quantity",
    values_to = "value"
  ) |>
  group_by(model, quantity) |>
  median_qi(value, .width = c(.8, .95)) |>
  tt()
```

# Summary of the socialist analysis

The socialist analysis uses the same estimands and modelling sequence as the radical-right analysis:

1. **Loyalty:** \(P(\text{socialist}\mid\text{high education})\)
2. **Low-education socialist support:** \(P(\text{socialist}\mid\text{low education})\)
3. **Education gap:** \(P(\text{socialist}\mid\text{high education})-P(\text{socialist}\mid\text{low education})\)
4. **Contribution:** \(P(\text{high education}\mid\text{socialist})\)

The main prior adjustment is to the intercept. The socialist-vote intercept is centred substantially higher than the radical-right intercept because the socialist outcome is more prevalent. The priors on the education coefficient, country-level standard deviations, correlations, and compositional covariates retain the same basic structure unless the prior predictive checks indicate that further adjustment is warranted.
