---
title: "Voting for the Radical Right in Germany: A First Bayesian Model"
subtitle: "Pr(rightvote | worker) — priors, prior predictive checks, and a simple hierarchical-free logistic model"
author:
  - name: Armin Schäfer
    affiliation:
      - id: jgu
        name: Johannes Gutenberg University
format:
  html:
    number-sections: true
    code-fold: true
    code-summary: "Show the code"
    self-contained: true
    toc: true
highlight-style: tango
execute:
  warning: false
  message: false
---

::: {.callout-note}
**Status: superseded by the pipeline.** After checking the fitted model
against this notebook's specification, the sample was narrowed further to
the 2017 and 2021 elections only, and to voters only (dropping non-voters
rather than folding them into "not radical right"). That finalized
specification now lives in `R/functions_clean.R`
(`clean_ess_ger_simple()`) and `R/functions_model.R`
(`fit_model_ger_simple()`), run via `_targets.R` (`targets::tar_make()`,
then `targets::tar_read(model_ger_simple)`). This notebook is kept as the
record of *how* the priors were derived and checked (still valid) and of
the exploratory `btw >= 2013`, population-based sample this model started
from (superseded, not wrong) — it is not the canonical version of the
model going forward.
:::

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
}

p_load(here, tidyverse, brms, tidybayes, ggdist, posterior, bayesplot, loo, marginaleffects, gt)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  dplyr.summarise.inform = FALSE,
  mc.cores = parallel::detectCores()
)

my_theme <- function(...) {
  theme(
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(color = "black", face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "grey70"),
    strip.text = element_text(color = "black"),
    legend.key.size = unit(0.3, "cm")
  )
}
```

# Purpose and scope

This notebook develops the *first, deliberately simple* Bayesian model in the
"from Bayes' rule to Bayesian estimation" workflow (see `ANALYSIS_PLAN.md`):
a single-country (Germany), non-hierarchical logistic regression of voting
for the radical right on worker status, with age, gender, immigrant
background, and union membership as controls. The substantive focus is the
posterior for `Pr(rightvote | worker)` versus `Pr(rightvote | non-worker)` —
the model-based, covariate-adjusted analogue of the manuscript's closed-form
Alford Index (`icv()` in `R/functions_derive.R`).

This is a **descriptive/associational** model: `worker` is an observed
occupational-class status, not a randomized treatment, so the coefficient on
`worker` should be read as an adjusted association net of age, gender,
immigrant background, and union membership — not a causal effect of "being a
worker" on vote choice.

Recoding logic here is exploratory (per `scripts/` convention); once
settled, it should move into `R/functions_clean.R` (`clean_ess()`) and
`R/functions_model.R` (replacing the `fit_model_pooled()` stub), and this
model becomes a `_targets.R` target rather than living only in this
notebook.

# Data and variable construction

```{r}
#| label: load-data

ess <- readRDS(here("data", "processed", "essger.rds"))
nrow(ess)

```

## Outcome: `rightvote`

The cleaned data provide `pvote_non_voter`, a party-choice variable that
already recodes non-voters into their own `"nonvoter"` category rather than
leaving them as `NA` (as plain `pvote` does). We use this population-based
coding — matching the manuscript's own framing of loyalty/contribution as
shares of the *population*, not just of voters — so `rightvote` compares
radical-right voters against everyone else, voters and non-voters alike. A
voters-only version of this model (dropping non-voters from the sample) is a
natural robustness check but is not the primary specification here.

```{r}
#| label: construct-rightvote

ess <- ess %>%
  mutate(
    rightvote = case_when(
      is.na(pvote_non_voter) ~ NA_real_,
      as.character(pvote_non_voter) == "radical right" ~ 1,
      TRUE ~ 0
    )
  )

count(ess, rightvote)
```

## Focal predictor: `worker`

`class5` is a five-category class scheme (Higher-grade service class /
Lower-grade service class / Small business / Skilled workers / Low-skilled
workers) derived upstream in `data/processed/class_scheme.rds` from detailed
occupational coding. Following Alford's (and Oesch's) definition of the
manual working class, `worker` combines the two manual-worker categories:

```{r}
#| label: construct-worker

ess <- ess %>%
  mutate(
    worker = case_when(
      is.na(class5) ~ NA_real_,
      class5 %in% c(4, 5) ~ 1,   # Skilled workers, Low-skilled workers
      TRUE ~ 0                    # Higher/Lower-grade service class, Small business
    )
  )

count(ess, worker, class5fac)
```

## A critical restriction: pre-2013 elections had (almost) no radical-right vote

The AfD was founded in 2013 and did not enter the Bundestag until 2017;
earlier German radical-right parties (Republikaner, NPD, DVU) polled far
below the threshold this coding scheme picks up. Pooling all eleven ESS
rounds (2002–2021) would silently mix elections in which "voting radical
right" was barely a live option with elections in which it was a major
choice — an identifying assumption that must be made explicit, not buried in
a pooled average:

```{r}
#| label: check-pre-afd

ess %>%
  group_by(btw) %>%
  summarise(n = n(), pct_rightvote = round(100 * mean(rightvote, na.rm = TRUE), 2))
```

Rightvote is exactly (or essentially) zero for 2002–2009. We therefore
restrict this first model to the AfD-era elections, 2013 onward. This drops
roughly half the sample but avoids building a model on three elections in
which the outcome was structurally impossible.

```{r}
#| label: restrict-sample

ess_afd_era <- ess %>% filter(btw >= 2013)
nrow(ess_afd_era)
```

## Reference categories and missingness

```{r}
#| label: recode-factors

ess_afd_era <- ess_afd_era %>%
  mutate(
    age_grp = factor(age_grp, levels = c("18-29", "30-65", "66+")),
    gender  = factor(gender,  levels = c("female", "male"))
  )

analysis_vars <- c("rightvote", "worker", "age_grp", "gender", "immigrant", "union")

n_before <- nrow(ess_afd_era)
ess_model <- ess_afd_era %>%
  drop_na(all_of(analysis_vars))
n_after <- nrow(ess_model)

tibble(
  n_before_listwise_deletion = n_before,
  n_after = n_after,
  pct_dropped = round(100 * (1 - n_after / n_before), 1)
)
```

Reference categories: `worker = 0` (non-worker), `age_grp = "18-29"`,
`gender = "female"`, `immigrant = 0`, `union = 0`. Missingness is
concentrated in `class5`/`worker` (occupational class is unavailable for
respondents outside the labour force); listwise deletion is a simplification
worth revisiting with multiple imputation once the model specification is
settled, but is acceptable for this first pass.

# Exploratory look at the focal contrast

```{r}
#| label: fig-eda-worker
#| fig-cap: "Radical-right vote share by worker status, Germany 2013-2021 (raw sample proportions, unadjusted for other covariates)."

eda_worker <- ess_model %>%
  group_by(worker) %>%
  summarise(
    n = n(),
    p_rightvote = mean(rightvote),
    se = sqrt(p_rightvote * (1 - p_rightvote) / n)
  ) %>%
  mutate(worker_lab = factor(worker, levels = c(0, 1), labels = c("Non-worker", "Worker")))

ggplot(eda_worker, aes(x = worker_lab, y = p_rightvote)) +
  geom_col(width = 0.5, fill = "grey40") +
  geom_errorbar(aes(ymin = p_rightvote - 1.96 * se, ymax = p_rightvote + 1.96 * se), width = 0.1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = "Radical-right vote share",
    title = "Raw (unadjusted) worker gap in radical-right voting"
  ) +
  my_theme()
```

This raw gap (roughly 3.5% vs. 6.5%, i.e. an unadjusted difference of about
3 percentage points, or an odds ratio of about 1.9) is the empirical
benchmark the priors below are checked against — priors are not fit to this
number, but should not rule it out a priori.

# The model

$$
\text{rightvote}_i \sim \text{Bernoulli}(p_i)
$$
$$
\text{logit}(p_i) = \alpha + \beta_{\text{worker}} \cdot \text{worker}_i +
\beta_{\text{age30-65}} \cdot \text{age30-65}_i + \beta_{\text{age66+}} \cdot \text{age66+}_i +
\beta_{\text{male}} \cdot \text{male}_i + \beta_{\text{immigrant}} \cdot \text{immigrant}_i +
\beta_{\text{union}} \cdot \text{union}_i
$$

No group-level (multilevel) structure is included at this stage: single
country, and the goal is a deliberately simple baseline before adding
country/round structure (see `ANALYSIS_PLAN.md`, models M1–M2). We use
brms's `0 + Intercept` parameterization so that `alpha` is the literal
log-odds for the reference respondent (young, non-worker, female,
non-immigrant, non-union) rather than brms's default internally-centered
intercept — this keeps the Intercept prior directly interpretable.

## Priors

| Parameter | Prior | Rationale |
|---|---|---|
| `alpha` (Intercept, reference respondent) | `Normal(-3, 1)` | `logit(-3) ≈ 4.7%`. AfD's actual national vote shares in 2013/2017/2021 were 4.7% / 12.6% / 10.3%; a young, non-worker, female, non-immigrant, non-union respondent is on the *low* end of that range on every dimension the model estimates, so centering below the national average is deliberate. SD = 1 puts 95% prior mass on `Pr(rightvote)` between 0.7% and 27% for this reference group — wide enough not to fight the data, tight enough to rule out the model treating >50% as remotely plausible a priori (well under 1% of prior mass, checked below). |
| `beta_worker` | `Normal(0, 0.5)` | Class-voting effects in the literature are moderate, not extreme. 95% prior mass on the odds ratio is roughly [0.37, 2.7] — allows the raw ~1.9 odds ratio above comfortably without presupposing it. |
| `beta_age30-65`, `beta_age66+` | `Normal(0, 0.5)` | Same weakly-informative logic; no specific directional expectation for age net of the other covariates. |
| `beta_male` | `Normal(0, 0.5)` | The gender gap in radical-right voting is one of the most robust findings in this literature and is likely to be at least as large as the worker gap; the same weakly-informative prior is used here deliberately, so that if the posterior for `beta_male` ends up needing to move further than for `beta_worker`, that is the *data* saying so, not the prior. |
| `beta_immigrant`, `beta_union` | `Normal(0, 0.5)` | Same weakly-informative default; the raw data show near-null unadjusted gaps for both, which this prior does not rule out. |

```{r}
#| label: set-priors

model_formula <- bf(rightvote ~ 0 + Intercept + worker + age_grp + gender + immigrant + union)

model_priors <- c(
  prior(normal(-3, 1), class = "b", coef = "Intercept"),
  prior(normal(0, 0.5), class = "b")
)
```

## Prior predictive check

Before fitting to real outcomes, two checks: a fast analytic simulation
(no Stan compilation needed, useful for quickly trying alternative prior
widths), then the canonical brms `sample_prior = "only"` check against the
actual model/data structure.

### Analytic check

```{r}
#| label: prior-predictive-analytic

set.seed(2026)
n_sim <- 20000

alpha     <- rnorm(n_sim, -3, 1)
b_worker  <- rnorm(n_sim, 0, 0.5)
b_age3065 <- rnorm(n_sim, 0, 0.5)
b_age66p  <- rnorm(n_sim, 0, 0.5)
b_male    <- rnorm(n_sim, 0, 0.5)
b_immig   <- rnorm(n_sim, 0, 0.5)
b_union   <- rnorm(n_sim, 0, 0.5)

p_reference   <- plogis(alpha)
p_worker      <- plogis(alpha + b_worker)
p_max_combo   <- plogis(alpha + b_worker + b_age3065 + b_male + b_immig + b_union)
icv_prior     <- p_worker - p_reference

qs <- function(x) round(quantile(x, c(.025, .25, .5, .75, .975)), 4)

tibble(
  quantity = c(
    "Pr(rightvote) at reference cell",
    "Pr(rightvote | worker=1, else reference)",
    "Pr(rightvote) at a maximal covariate combination",
    "Prior-predictive worker contrast (ICV-like)"
  )
) %>%
  bind_cols(
    map_dfr(list(p_reference, p_worker, p_max_combo, icv_prior), ~ as_tibble_row(qs(.x)))
  ) %>%
  gt() %>%
  fmt_number(columns = -quantity, decimals = 3)

cat("P(reference-cell Pr(rightvote) > 0.5):", mean(p_reference > 0.5), "\n")
cat("P(maximal-combination Pr(rightvote) > 0.5):", mean(p_max_combo > 0.5), "\n")
cat("P(prior-predictive worker contrast > 0):", mean(icv_prior > 0), "\n")
```

The prior-predictive worker contrast is centered at essentially zero
(`P(contrast > 0) ≈ 0.5`) with a 95% interval comfortably spanning the raw
sample gap (~+3 points) in both directions — the prior is not pre-judging
the sign or size of the worker effect. Implausible predictions (>50% radical
right vote share) get well under 3% of prior mass even at the most
AfD-favorable covariate combination.

### Canonical brms check

```{r}
#| label: prior-predictive-brms
#| eval: false

fit_prior_only <- brm(
  formula = model_formula,
  data    = ess_model,
  family  = bernoulli(),
  prior   = model_priors,
  sample_prior = "only",
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_simple_prior_only")
)

pp_check(fit_prior_only, ndraws = 100) +
  labs(title = "Prior predictive check: rightvote ~ worker + age_grp + gender + immigrant + union")
```

`eval: false` above only because this notebook was drafted in an environment
without a working Stan toolchain to verify the MCMC step end-to-end (the
data construction, priors, and analytic check above *were* run against the
real, staged data). Set to `eval: true` (or delete the option) before
running locally — inspect the plot for implausible mass before proceeding to
the real fit.

## Fitting the model

```{r}
#| label: fit-model
#| eval: false

fit <- brm(
  formula = model_formula,
  data    = ess_model,
  family  = bernoulli(),
  prior   = model_priors,
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_simple")
)
```

## Diagnostics

```{r}
#| label: diagnostics
#| eval: false

summary(fit)
posterior::summarise_draws(fit, "rhat", "ess_bulk", "ess_tail")

pp_check(fit, ndraws = 100)
pp_check(fit, type = "stat_grouped", group = "worker", stat = "mean")
```

Check before interpreting anything below: all `Rhat < 1.01`, bulk- and
tail-ESS comfortably above ~400 per parameter, and the grouped posterior
predictive check shows the model reproducing the observed radical-right vote
share within each worker group reasonably well. Do not proceed past this
point on the basis of this notebook's code alone — inspect the actual
numbers and plots yourself.

# Focal result: Pr(rightvote | worker)

```{r}
#| label: focal-contrast
#| eval: false

worker_contrast <- avg_predictions(
  fit,
  variables = "worker",
  type = "response"
)
worker_contrast

# `add_epred_draws()` builds its prediction from every population-level
# predictor in the model, so `newdata` must supply all of them -- unlike
# `re_formula = NA` (which only drops *group-level* terms, none of which
# exist in this flat model), there is no way to omit age_grp/gender/
# immigrant/union here. We hold them at their reference levels explicitly,
# including the *unused* factor levels (via `levels = levels(ess_model$...)`)
# so brms reconstructs the same dummy-coded columns it was fitted on --
# a factor built from a single string without this would silently produce
# a 1-level factor and break the model matrix.
newdata_worker <- tibble(
  worker    = c(0, 1),
  age_grp   = factor("18-29", levels = levels(ess_model$age_grp)),
  gender    = factor("female", levels = levels(ess_model$gender)),
  immigrant = 0,
  union     = 0
)

worker_draws <- newdata_worker %>%
  add_epred_draws(fit, re_formula = NA) %>%
  ungroup()

worker_draws %>%
  mutate(worker_lab = factor(worker, levels = c(0, 1), labels = c("Non-worker", "Worker"))) %>%
  ggplot(aes(x = .epred, y = worker_lab, fill = worker_lab)) +
  stat_halfeye(alpha = 0.7) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_viridis_d(guide = "none") +
  labs(
    x = "Posterior Pr(rightvote)", y = NULL,
    title = "Posterior distribution of Pr(rightvote | worker), other covariates at reference levels"
  ) +
  my_theme()

posterior_icv <- worker_draws %>%
  select(.draw, worker, .epred) %>%
  pivot_wider(names_from = worker, values_from = .epred, names_prefix = "p_worker") %>%
  mutate(posterior_icv = p_worker1 - p_worker0)

posterior_icv %>%
  summarise(
    median = median(posterior_icv),
    q025 = quantile(posterior_icv, .025),
    q975 = quantile(posterior_icv, .975),
    p_gt_0 = mean(posterior_icv > 0)
  )
```

`worker_contrast` and `posterior_icv` answer a related but distinct
question and should not be reported interchangeably: `worker_contrast`
(via `avg_predictions()`) is the *average marginal effect* of `worker`,
averaged over the observed sample distribution of age/gender/immigrant/
union; `posterior_icv` is the effect for one specific, fixed covariate
profile (young, female, non-immigrant, non-union). They will normally be
close but need not be identical, particularly if `worker` is unevenly
distributed across the other covariates (e.g. if workers skew male).

`posterior_icv` is the covariate-adjusted, fully Bayesian analogue of the
manuscript's closed-form `icv()`: instead of one point estimate from a raw
2x2 table, it is a full posterior distribution of the worker/non-worker gap
in `Pr(rightvote)`, net of age, gender, immigrant background, and union
membership. Report it as median + 95% credible interval, and note
`p_gt_0` — the posterior probability that the adjusted worker effect is
positive — rather than reducing it to a single point estimate with a
p-value.

# Next steps for this model

- [x] ~~Restrict to voters only.~~ Done: `clean_ess_ger_simple()` drops
  non-voters (`pvote == NA`) instead of folding them into "not radical
  right".
- [x] ~~Once satisfied with the specification, move `rightvote`/`worker`
  construction into `R/functions_clean.R` and this model into
  `R/functions_model.R`, and add both as `_targets.R` targets.~~ Done —
  see the status note at the top of this notebook. The sample was also
  narrowed from `btw >= 2013` to `btw %in% c(2017, 2021)` in the process.
- Run `targets::tar_make()` (Stan toolchain required for `model_ger_simple`)
  and inspect `targets::tar_read(model_ger_simple)`'s diagnostics
  (`summary()`, `pp_check()`, Rhat/ESS) before trusting any posterior
  result from it.
- Sensitivity check: refit with `beta_worker ~ Normal(0, 0.25)` and
  `Normal(0, 1)` and confirm the posterior conclusion about `p_gt_0` is not
  an artifact of the SD = 0.5 choice (ANALYSIS_PLAN.md, §7) — a good use of
  `tar_target()`'s branching, or just a manual second call to
  `fit_model_ger_simple()` with an adjusted prior argument.
- Extend to the multilevel model sequence in `ANALYSIS_PLAN.md` (M1/M2) once
  ready to move beyond a single country.
- Consider formalizing `scripts/process_ess_ger*.qmd` (which build
  `essger.rds` from raw ESS + the class scheme) into `R/` functions and
  targets too, so the whole pipeline — not just this model — is
  `tar_make()`-reproducible from raw ESS data.

# Reverse model: Pr(worker | rightvote)

The model above answers the *voter perspective* — Alford's ICV question,
"are workers disproportionately likely to vote radical right?" The
manuscript's other half is the *party perspective* — Axelrod's Contribution
Index / Hooghe & Marks' PCI question, "what share of the radical-right
electorate is made up of workers?" These are two sides of the same Bayes'
rule identity (see `R/functions_derive.R::contribution()`), but rather than
deriving one from the other analytically — which would require also
modeling `Pr(rightvote)` and `Pr(worker)` as separate quantities with their
own uncertainty — this fits the reverse regression directly: `worker` as
the outcome, `rightvote` as the predictor, on its own likelihood. Both are
equally legitimate "Bayesian models" of the same underlying joint
distribution; fitting directly is simpler here because there is only one
predictor and no adjustment question in view yet.

Deliberately excludes age_grp/gender/immigrant/union: the goal stated for
this model is the *unconditional* share of workers among radical-right
voters (the direct analogue of Axelrod's and Hooghe & Marks' definitions,
which are not covariate-adjusted either), not a demographics-adjusted
version of it. Adding those controls back in would answer a different,
also useful question — "net of age/gender/immigrant background/union
membership, how much of the raw worker-share gap is a residual class
effect?" — flagged as a next step, not done here.

## Data

Reuses the exact same finalized analysis sample as the model above (same
`clean_ess_ger_simple()`, same N = 7,889): Germany, 2017 + 2021 Bundestag
elections, voters only.

```{r}
#| label: reverse-data

# Re-read essger.rds fresh rather than reusing `ess` from the sections
# above: by this point `ess` carries the superseded rightvote/worker
# columns from the btw>=2013, population-based exploration. clean_ess_ger_simple()
# would overwrite them correctly either way, but starting fresh avoids any
# doubt about which definition is in effect.
source(here("R", "functions_clean.R"))
essger_raw    <- readRDS(here("data", "processed", "essger.rds"))
ess_model_rev <- clean_ess_ger_simple(essger_raw)
nrow(ess_model_rev)
```

## Exploratory look at the focal contrast

```{r}
#| label: fig-eda-reverse
#| fig-cap: "Worker share by vote choice, Germany 2017+2021 voters (raw sample proportions)."

eda_reverse <- ess_model_rev %>%
  group_by(rightvote) %>%
  summarise(
    n = n(),
    p_worker = mean(worker),
    se = sqrt(p_worker * (1 - p_worker) / n)
  ) %>%
  mutate(rightvote_lab = factor(rightvote, levels = c(0, 1), labels = c("Other/no radical-right vote", "Radical-right vote")))

ggplot(eda_reverse, aes(x = rightvote_lab, y = p_worker)) +
  geom_col(width = 0.5, fill = "grey40") +
  geom_errorbar(aes(ymin = p_worker - 1.96 * se, ymax = p_worker + 1.96 * se), width = 0.1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = "Worker share",
    title = "Raw (unadjusted) worker-share gap by vote choice"
  ) +
  my_theme()
```

The raw gap — roughly 38% among other voters vs. 59% among radical-right
voters, an odds ratio of about 2.3 — is unusually large for a single
demographic predictor, but is consistent with how strongly class-based the
literature on West European radical-right electorates finds this
particular divide to be; it is the benchmark the priors below are checked
against, not fit to.

## The model

$$
\text{worker}_i \sim \text{Bernoulli}(p_i)
$$
$$
\text{logit}(p_i) = \alpha + \beta_{\text{rightvote}} \cdot \text{rightvote}_i
$$

Same `0 + Intercept` parameterization as the forward model, so `alpha` is
the literal log-odds of being a worker among non-radical-right voters (the
reference level), not brms's internally-centered intercept.

### Priors

| Parameter | Prior | Rationale |
|---|---|---|
| `alpha` (Intercept, worker share among non-RR voters) | `Normal(-0.3, 1)` | `logit(-0.3) ≈ 43%`. Workers are a large minority, not a majority, of contemporary West European electorates under class schemes like the one used here (consistent with this project's own earlier, broader-sample worker-share estimates of ~44-50% in `scripts/analyze_ess_ger.qmd`'s exploratory data, though that is a different subsample from the one used to fit this model, so this is a plausibility check rather than an empirical-Bayes prior). SD = 1 puts 95% prior mass on `Pr(worker \| rightvote=0)` between 10% and 84% — wide, deliberately not fit to this specific sample's 37.7%. |
| `beta_rightvote` | `Normal(0, 0.75)` | Wider than the demographic-control priors in the forward model (`Normal(0, 0.5)`) because this coefficient is not a nuisance control here — it is the very relationship the model exists to estimate, and the class divide in radical-right electorates is one of the more robust *large* effects in this literature (not a small nudge like the immigrant/union coefficients in the forward model). 95% prior mass on the odds ratio is roughly [0.23, 4.3], comfortably covering the raw sample OR of ~2.3 without presupposing it (`P(contrast > 0) = 0.50` under the prior, checked below). |

```{r}
#| label: reverse-set-priors

reverse_formula <- bf(worker ~ 0 + Intercept + rightvote)

reverse_priors <- c(
  prior(normal(-0.3, 1), class = "b", coef = "Intercept"),
  prior(normal(0, 0.75), class = "b", coef = "rightvote")
)
```

### Prior predictive check (analytic)

```{r}
#| label: reverse-prior-predictive-analytic

set.seed(2026)
n_sim <- 20000
alpha       <- rnorm(n_sim, -0.3, 1)
b_rightvote <- rnorm(n_sim, 0, 0.75)

p_nonrv  <- plogis(alpha)
p_rv     <- plogis(alpha + b_rightvote)
contrast <- p_rv - p_nonrv

qs <- function(x) round(quantile(x, c(.025, .25, .5, .75, .975)), 4)

tibble(quantity = c(
  "Pr(worker | rightvote=0)",
  "Pr(worker | rightvote=1) -- the requested quantity",
  "Contrast: Pr(worker|rightvote=1) - Pr(worker|rightvote=0)"
)) %>%
  bind_cols(map_dfr(list(p_nonrv, p_rv, contrast), ~ as_tibble_row(qs(.x)))) %>%
  gt() %>%
  fmt_number(columns = -quantity, decimals = 3)

cat("P(contrast > 0):", mean(contrast > 0), "\n")
cat("P(Pr(worker|rightvote=1) > 0.95):", mean(p_rv > 0.95), "\n")
cat("P(Pr(worker|rightvote=1) < 0.05):", mean(p_rv < 0.05), "\n")
```

The prior-predictive contrast is centered at 0 (not presupposing a
direction) with a 95% interval of roughly [-28, +30] percentage points,
comfortably bracketing the raw sample gap of +21 points; implausible
extremes (`Pr(worker) > 95%` or `< 5%`) get well under 2% of prior mass
each.

### Canonical brms check and fit

```{r}
#| label: reverse-prior-predictive-brms
#| eval: false

fit_rev_prior_only <- brm(
  formula = reverse_formula,
  data    = ess_model_rev,
  family  = bernoulli(),
  prior   = reverse_priors,
  sample_prior = "only",
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_reverse_prior_only")
)

pp_check(fit_rev_prior_only, ndraws = 100) +
  labs(title = "Prior predictive check: worker ~ rightvote")
```

```{r}
#| label: reverse-fit-model
#| eval: false

fit_rev <- brm(
  formula = reverse_formula,
  data    = ess_model_rev,
  family  = bernoulli(),
  prior   = reverse_priors,
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_reverse")
)
```

### Diagnostics

```{r}
#| label: reverse-diagnostics
#| eval: false

summary(fit_rev)
posterior::summarise_draws(fit_rev, "rhat", "ess_bulk", "ess_tail")

pp_check(fit_rev, ndraws = 100)
pp_check(fit_rev, type = "stat_grouped", group = "rightvote", stat = "mean")
```

As with the forward model: check `Rhat < 1.01` and bulk/tail ESS before
trusting anything below, and inspect the grouped posterior predictive check
yourself rather than taking this notebook's word for it.

## Focal result: the share of workers among radical-right voters

```{r}
#| label: reverse-focal-result
#| eval: false

worker_share_draws <- tibble(rightvote = c(0, 1)) %>%
  add_epred_draws(fit_rev) %>%
  ungroup()

worker_share_draws %>%
  mutate(rightvote_lab = factor(rightvote, levels = c(0, 1), labels = c("Other/no radical-right vote", "Radical-right vote"))) %>%
  ggplot(aes(x = .epred, y = rightvote_lab, fill = rightvote_lab)) +
  stat_halfeye(alpha = 0.7) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_viridis_d(guide = "none") +
  labs(
    x = "Posterior Pr(worker)", y = NULL,
    title = "Posterior share of workers, by vote choice"
  ) +
  my_theme()

worker_share_among_rr <- worker_share_draws %>%
  filter(rightvote == 1) %>%
  summarise(
    median = median(.epred),
    q025 = quantile(.epred, .025),
    q975 = quantile(.epred, .975)
  )
worker_share_among_rr
```

`worker_share_among_rr` is the answer to the question this section set out
to answer: the posterior-median predicted share of workers among
radical-right voters, with a 95% credible interval — report it that way
(e.g. "an estimated X% [95% CI: Y-Z%] of radical-right voters are workers"),
not as a single point estimate. This is the fully Bayesian analogue of
Axelrod's Contribution Index / Hooghe & Marks' PCI's core quantity
(`Pr(worker | party)`); subtracting the population worker share (`Pr(worker)`,
~38.9% in this sample) from it recovers a PCI-like differential, though a
fully rigorous posterior for that subtraction should also propagate
`Pr(worker)`'s own sampling uncertainty rather than treating it as a fixed
number — left as a next step, not done here.

# Next steps for the reverse model

- Run `targets::tar_make()`-style fitting locally (Stan toolchain required)
  and replace the `eval: false` chunks' placeholders with real posterior
  output.
- Once satisfied, promote `worker ~ rightvote` the same way the forward
  model was promoted: a `clean` step is already shared
  (`clean_ess_ger_simple()`), so this mainly means adding
  `fit_model_ger_reverse()` to `R/functions_model.R` and a
  `model_ger_reverse` target to `_targets.R`.
- Decide whether to build the full posterior PCI (`Pr(worker|rightvote) -
  Pr(worker)`, both terms with their own posterior uncertainty, combined
  draw-by-draw) rather than subtracting a fixed population share.
- Consider adding age_grp/gender/immigrant/union back in to decompose how
  much of the raw worker-share gap survives adjustment for demographic
  composition, once ready to ask that (different, adjusted) question.
