---
title: "Voting for the Radical Right in Germany: A First Bayesian Model"
subtitle: "Pr(rightvote | worker) — priors, prior predictive checks, and a simple hierarchical-free logistic model"
author:
  - name: Armin Schäfer
    affiliation:
      - id: jgu
        name: Johannes Gutenberg University
format:
  html:
    number-sections: true
    code-fold: true
    code-summary: "Show the code"
    self-contained: true
    toc: true
highlight-style: tango
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
}

p_load(here, tidyverse, brms, tidybayes, ggdist, posterior, bayesplot, loo, marginaleffects, gt)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  dplyr.summarise.inform = FALSE,
  mc.cores = parallel::detectCores()
)

my_theme <- function(...) {
  theme(
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(color = "black", face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white", color = "grey70"),
    strip.text = element_text(color = "black"),
    legend.key.size = unit(0.3, "cm")
  )
}
```

# Purpose and scope

This notebook develops the *first, deliberately simple* Bayesian model in the
"from Bayes' rule to Bayesian estimation" workflow (see `ANALYSIS_PLAN.md`):
a single-country (Germany), non-hierarchical logistic regression of voting
for the radical right on worker status, with age, gender, immigrant
background, and union membership as controls. The substantive focus is the
posterior for `Pr(rightvote | worker)` versus `Pr(rightvote | non-worker)` —
the model-based, covariate-adjusted analogue of the manuscript's closed-form
Alford Index (`icv()` in `R/functions_derive.R`).

This is a **descriptive/associational** model: `worker` is an observed
occupational-class status, not a randomized treatment, so the coefficient on
`worker` should be read as an adjusted association net of age, gender,
immigrant background, and union membership — not a causal effect of "being a
worker" on vote choice.

Recoding logic here is exploratory (per `scripts/` convention); once
settled, it should move into `R/functions_clean.R` (`clean_ess()`) and
`R/functions_model.R` (replacing the `fit_model_pooled()` stub), and this
model becomes a `_targets.R` target rather than living only in this
notebook.

# Data and variable construction

```{r}
#| label: load-data

ess <- readRDS(here("data", "processed", "essger.rds"))
nrow(ess)
```

## Outcome: `rightvote`

The cleaned data provide `pvote_non_voter`, a party-choice variable that
already recodes non-voters into their own `"nonvoter"` category rather than
leaving them as `NA` (as plain `pvote` does). We use this population-based
coding — matching the manuscript's own framing of loyalty/contribution as
shares of the *population*, not just of voters — so `rightvote` compares
radical-right voters against everyone else, voters and non-voters alike. A
voters-only version of this model (dropping non-voters from the sample) is a
natural robustness check but is not the primary specification here.

```{r}
#| label: construct-rightvote

ess <- ess %>%
  mutate(
    rightvote = case_when(
      is.na(pvote_non_voter) ~ NA_real_,
      as.character(pvote_non_voter) == "radical right" ~ 1,
      TRUE ~ 0
    )
  )

count(ess, rightvote)
```

## Focal predictor: `worker`

`class5` is a five-category class scheme (Higher-grade service class /
Lower-grade service class / Small business / Skilled workers / Low-skilled
workers) derived upstream in `data/processed/class_scheme.rds` from detailed
occupational coding. Following Alford's (and Oesch's) definition of the
manual working class, `worker` combines the two manual-worker categories:

```{r}
#| label: construct-worker

ess <- ess %>%
  mutate(
    worker = case_when(
      is.na(class5) ~ NA_real_,
      class5 %in% c(4, 5) ~ 1,   # Skilled workers, Low-skilled workers
      TRUE ~ 0                    # Higher/Lower-grade service class, Small business
    )
  )

count(ess, worker, class5fac)
```

## A critical restriction: pre-2013 elections had (almost) no radical-right vote

The AfD was founded in 2013 and did not enter the Bundestag until 2017;
earlier German radical-right parties (Republikaner, NPD, DVU) polled far
below the threshold this coding scheme picks up. Pooling all eleven ESS
rounds (2002–2021) would silently mix elections in which "voting radical
right" was barely a live option with elections in which it was a major
choice — an identifying assumption that must be made explicit, not buried in
a pooled average:

```{r}
#| label: check-pre-afd

ess %>%
  group_by(btw) %>%
  summarise(n = n(), pct_rightvote = round(100 * mean(rightvote, na.rm = TRUE), 2))
```

Rightvote is exactly (or essentially) zero for 2002–2009. We therefore
restrict this first model to the AfD-era elections, 2013 onward. This drops
roughly half the sample but avoids building a model on three elections in
which the outcome was structurally impossible.

```{r}
#| label: restrict-sample

ess_afd_era <- ess %>% filter(btw >= 2013)
nrow(ess_afd_era)
```

## Reference categories and missingness

```{r}
#| label: recode-factors

ess_afd_era <- ess_afd_era %>%
  mutate(
    age_grp = factor(age_grp, levels = c("18-29", "30-65", "66+")),
    gender  = factor(gender,  levels = c("female", "male"))
  )

analysis_vars <- c("rightvote", "worker", "age_grp", "gender", "immigrant", "union")

n_before <- nrow(ess_afd_era)
ess_model <- ess_afd_era %>%
  drop_na(all_of(analysis_vars))
n_after <- nrow(ess_model)

tibble(
  n_before_listwise_deletion = n_before,
  n_after = n_after,
  pct_dropped = round(100 * (1 - n_after / n_before), 1)
)
```

Reference categories: `worker = 0` (non-worker), `age_grp = "18-29"`,
`gender = "female"`, `immigrant = 0`, `union = 0`. Missingness is
concentrated in `class5`/`worker` (occupational class is unavailable for
respondents outside the labour force); listwise deletion is a simplification
worth revisiting with multiple imputation once the model specification is
settled, but is acceptable for this first pass.

# Exploratory look at the focal contrast

```{r}
#| label: fig-eda-worker
#| fig-cap: "Radical-right vote share by worker status, Germany 2013-2021 (raw sample proportions, unadjusted for other covariates)."

eda_worker <- ess_model %>%
  group_by(worker) %>%
  summarise(
    n = n(),
    p_rightvote = mean(rightvote),
    se = sqrt(p_rightvote * (1 - p_rightvote) / n)
  ) %>%
  mutate(worker_lab = factor(worker, levels = c(0, 1), labels = c("Non-worker", "Worker")))

ggplot(eda_worker, aes(x = worker_lab, y = p_rightvote)) +
  geom_col(width = 0.5, fill = "grey40") +
  geom_errorbar(aes(ymin = p_rightvote - 1.96 * se, ymax = p_rightvote + 1.96 * se), width = 0.1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = "Radical-right vote share",
    title = "Raw (unadjusted) worker gap in radical-right voting"
  ) +
  my_theme()
```

This raw gap (roughly 3.5% vs. 6.5%, i.e. an unadjusted difference of about
3 percentage points, or an odds ratio of about 1.9) is the empirical
benchmark the priors below are checked against — priors are not fit to this
number, but should not rule it out a priori.

# The model

$$
\text{rightvote}_i \sim \text{Bernoulli}(p_i)
$$
$$
\text{logit}(p_i) = \alpha + \beta_{\text{worker}} \cdot \text{worker}_i +
\beta_{\text{age30-65}} \cdot \text{age30-65}_i + \beta_{\text{age66+}} \cdot \text{age66+}_i +
\beta_{\text{male}} \cdot \text{male}_i + \beta_{\text{immigrant}} \cdot \text{immigrant}_i +
\beta_{\text{union}} \cdot \text{union}_i
$$

No group-level (multilevel) structure is included at this stage: single
country, and the goal is a deliberately simple baseline before adding
country/round structure (see `ANALYSIS_PLAN.md`, models M1–M2). We use
brms's `0 + Intercept` parameterization so that `alpha` is the literal
log-odds for the reference respondent (young, non-worker, female,
non-immigrant, non-union) rather than brms's default internally-centered
intercept — this keeps the Intercept prior directly interpretable.

## Priors

| Parameter | Prior | Rationale |
|---|---|---|
| `alpha` (Intercept, reference respondent) | `Normal(-3, 1)` | `logit(-3) ≈ 4.7%`. AfD's actual national vote shares in 2013/2017/2021 were 4.7% / 12.6% / 10.3%; a young, non-worker, female, non-immigrant, non-union respondent is on the *low* end of that range on every dimension the model estimates, so centering below the national average is deliberate. SD = 1 puts 95% prior mass on `Pr(rightvote)` between 0.7% and 27% for this reference group — wide enough not to fight the data, tight enough to rule out the model treating >50% as remotely plausible a priori (well under 1% of prior mass, checked below). |
| `beta_worker` | `Normal(0, 0.5)` | Class-voting effects in the literature are moderate, not extreme. 95% prior mass on the odds ratio is roughly [0.37, 2.7] — allows the raw ~1.9 odds ratio above comfortably without presupposing it. |
| `beta_age30-65`, `beta_age66+` | `Normal(0, 0.5)` | Same weakly-informative logic; no specific directional expectation for age net of the other covariates. |
| `beta_male` | `Normal(0, 0.5)` | The gender gap in radical-right voting is one of the most robust findings in this literature and is likely to be at least as large as the worker gap; the same weakly-informative prior is used here deliberately, so that if the posterior for `beta_male` ends up needing to move further than for `beta_worker`, that is the *data* saying so, not the prior. |
| `beta_immigrant`, `beta_union` | `Normal(0, 0.5)` | Same weakly-informative default; the raw data show near-null unadjusted gaps for both, which this prior does not rule out. |

```{r}
#| label: set-priors

model_formula <- bf(rightvote ~ 0 + Intercept + worker + age_grp + gender + immigrant + union)

model_priors <- c(
  prior(normal(-3, 1), class = "b", coef = "Intercept"),
  prior(normal(0, 0.5), class = "b")
)
```

## Prior predictive check

Before fitting to real outcomes, two checks: a fast analytic simulation
(no Stan compilation needed, useful for quickly trying alternative prior
widths), then the canonical brms `sample_prior = "only"` check against the
actual model/data structure.

### Analytic check

```{r}
#| label: prior-predictive-analytic

set.seed(2026)
n_sim <- 20000

alpha     <- rnorm(n_sim, -3, 1)
b_worker  <- rnorm(n_sim, 0, 0.5)
b_age3065 <- rnorm(n_sim, 0, 0.5)
b_age66p  <- rnorm(n_sim, 0, 0.5)
b_male    <- rnorm(n_sim, 0, 0.5)
b_immig   <- rnorm(n_sim, 0, 0.5)
b_union   <- rnorm(n_sim, 0, 0.5)

p_reference   <- plogis(alpha)
p_worker      <- plogis(alpha + b_worker)
p_max_combo   <- plogis(alpha + b_worker + b_age3065 + b_male + b_immig + b_union)
icv_prior     <- p_worker - p_reference

qs <- function(x) round(quantile(x, c(.025, .25, .5, .75, .975)), 4)

tibble(
  quantity = c(
    "Pr(rightvote) at reference cell",
    "Pr(rightvote | worker=1, else reference)",
    "Pr(rightvote) at a maximal covariate combination",
    "Prior-predictive worker contrast (ICV-like)"
  )
) %>%
  bind_cols(
    map_dfr(list(p_reference, p_worker, p_max_combo, icv_prior), ~ as_tibble_row(qs(.x)))
  ) %>%
  gt() %>%
  fmt_number(columns = -quantity, decimals = 3)

cat("P(reference-cell Pr(rightvote) > 0.5):", mean(p_reference > 0.5), "\n")
cat("P(maximal-combination Pr(rightvote) > 0.5):", mean(p_max_combo > 0.5), "\n")
cat("P(prior-predictive worker contrast > 0):", mean(icv_prior > 0), "\n")
```

The prior-predictive worker contrast is centered at essentially zero
(`P(contrast > 0) ≈ 0.5`) with a 95% interval comfortably spanning the raw
sample gap (~+3 points) in both directions — the prior is not pre-judging
the sign or size of the worker effect. Implausible predictions (>50% radical
right vote share) get well under 3% of prior mass even at the most
AfD-favorable covariate combination.

### Canonical brms check

```{r}
#| label: prior-predictive-brms
#| eval: false

fit_prior_only <- brm(
  formula = model_formula,
  data    = ess_model,
  family  = bernoulli(),
  prior   = model_priors,
  sample_prior = "only",
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_simple_prior_only")
)

pp_check(fit_prior_only, ndraws = 100) +
  labs(title = "Prior predictive check: rightvote ~ worker + age_grp + gender + immigrant + union")
```

`eval: false` above only because this notebook was drafted in an environment
without a working Stan toolchain to verify the MCMC step end-to-end (the
data construction, priors, and analytic check above *were* run against the
real, staged data). Set to `eval: true` (or delete the option) before
running locally — inspect the plot for implausible mass before proceeding to
the real fit.

## Fitting the model

```{r}
#| label: fit-model
#| eval: false

fit <- brm(
  formula = model_formula,
  data    = ess_model,
  family  = bernoulli(),
  prior   = model_priors,
  chains = 4, cores = 4, iter = 2000, seed = 2026,
  file = here("models", "m_ger_simple")
)
```

## Diagnostics

```{r}
#| label: diagnostics
#| eval: false

summary(fit)
posterior::summarise_draws(fit, "rhat", "ess_bulk", "ess_tail")

pp_check(fit, ndraws = 100)
pp_check(fit, type = "stat_grouped", group = "worker", stat = "mean")
```

Check before interpreting anything below: all `Rhat < 1.01`, bulk- and
tail-ESS comfortably above ~400 per parameter, and the grouped posterior
predictive check shows the model reproducing the observed radical-right vote
share within each worker group reasonably well. Do not proceed past this
point on the basis of this notebook's code alone — inspect the actual
numbers and plots yourself.

# Focal result: Pr(rightvote | worker)

```{r}
#| label: focal-contrast
#| eval: false

worker_contrast <- avg_predictions(
  fit,
  variables = "worker",
  type = "response"
)
worker_contrast

# `add_epred_draws()` builds its prediction from every population-level
# predictor in the model, so `newdata` must supply all of them -- unlike
# `re_formula = NA` (which only drops *group-level* terms, none of which
# exist in this flat model), there is no way to omit age_grp/gender/
# immigrant/union here. We hold them at their reference levels explicitly,
# including the *unused* factor levels (via `levels = levels(ess_model$...)`)
# so brms reconstructs the same dummy-coded columns it was fitted on --
# a factor built from a single string without this would silently produce
# a 1-level factor and break the model matrix.
newdata_worker <- tibble(
  worker    = c(0, 1),
  age_grp   = factor("18-29", levels = levels(ess_model$age_grp)),
  gender    = factor("female", levels = levels(ess_model$gender)),
  immigrant = 0,
  union     = 0
)

worker_draws <- newdata_worker %>%
  add_epred_draws(fit, re_formula = NA) %>%
  ungroup()

worker_draws %>%
  mutate(worker_lab = factor(worker, levels = c(0, 1), labels = c("Non-worker", "Worker"))) %>%
  ggplot(aes(x = .epred, y = worker_lab, fill = worker_lab)) +
  stat_halfeye(alpha = 0.7) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_viridis_d(guide = "none") +
  labs(
    x = "Posterior Pr(rightvote)", y = NULL,
    title = "Posterior distribution of Pr(rightvote | worker), other covariates at reference levels"
  ) +
  my_theme()

posterior_icv <- worker_draws %>%
  select(.draw, worker, .epred) %>%
  pivot_wider(names_from = worker, values_from = .epred, names_prefix = "p_worker") %>%
  mutate(posterior_icv = p_worker1 - p_worker0)

posterior_icv %>%
  summarise(
    median = median(posterior_icv),
    q025 = quantile(posterior_icv, .025),
    q975 = quantile(posterior_icv, .975),
    p_gt_0 = mean(posterior_icv > 0)
  )
```

`worker_contrast` and `posterior_icv` answer a related but distinct
question and should not be reported interchangeably: `worker_contrast`
(via `avg_predictions()`) is the *average marginal effect* of `worker`,
averaged over the observed sample distribution of age/gender/immigrant/
union; `posterior_icv` is the effect for one specific, fixed covariate
profile (young, female, non-immigrant, non-union). They will normally be
close but need not be identical, particularly if `worker` is unevenly
distributed across the other covariates (e.g. if workers skew male).

`posterior_icv` is the covariate-adjusted, fully Bayesian analogue of the
manuscript's closed-form `icv()`: instead of one point estimate from a raw
2x2 table, it is a full posterior distribution of the worker/non-worker gap
in `Pr(rightvote)`, net of age, gender, immigrant background, and union
membership. Report it as median + 95% credible interval, and note
`p_gt_0` — the posterior probability that the adjusted worker effect is
positive — rather than reducing it to a single point estimate with a
p-value.

# Next steps

- Run the two `eval: false` chunks locally (Stan toolchain required) and
  replace this section with the actual diagnostic output and posterior
  summaries.
- Sensitivity check: refit with `beta_worker ~ Normal(0, 0.25)` and
  `Normal(0, 1)` and confirm the posterior conclusion about `p_gt_0` is not
  an artifact of the SD = 0.5 choice (ANALYSIS_PLAN.md, §7).
- Robustness check: restrict to voters only (drop `nonvoter` from
  `pvote_non_voter`) and compare.
- Once satisfied with the specification, move `rightvote`/`worker`
  construction into `R/functions_clean.R` and this model into
  `R/functions_model.R`, and add both as `_targets.R` targets.
- Extend to the multilevel model sequence in `ANALYSIS_PLAN.md` (M1/M2) once
  ready to move beyond a single country.
