---
title: "ESS_EduXGenXGend_Recode"
author: "Research Assistant Lars Hinkel"
date: "`r Sys.Date()`"
output: 
  html_document: 
    df_print: paged
editor: 
  markdown: 
    wrap: sentence
---

# Load packages using Packman and pload

```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, haven, here, labelled, dplyr, readr, rio, forcats, tidyr)
```

# Load and preprocess data from ESS-Data-Wizard

```{r}

ess_filtered     <- import(here("data", "raw", "ESS_rounds_1-11.csv"), trust = TRUE)

ess_class_scheme <- import(here("data", "processed", "ess_class_scheme.rds"), trust = TRUE)

ess_filtered <- merge(ess_filtered, ess_class_scheme, by = c("essround", "idno"))


```

# Data preparation

## Filtering and generating macro IDs

```{r}
# Data preparation
# Filter: Only Western European countries included in ESS rounds 1-11
#ess_filtered <- ess |>
#  filter(cntry %in% c("BE", "CH", "DE", "ES", "FI", "FR", "GB", "IE", "NL", "NO", "PT", "SE"))

# Generate macro IDs
ess_filtered <- ess_filtered |>
  mutate(
    cntry_id = as.numeric(as.factor(cntry)),
    cntry_wave_id = as.numeric(as.factor(paste(cntry, essround, sep = "_")))
  )

# Sort data by `cntry` and `essround` before generating IDs
ess_filtered <- ess_filtered |>
  arrange(cntry, essround) |>
  mutate(
    cntry_id = as.numeric(factor(cntry, levels = unique(cntry))), # Takes data order into account
    cntry_wave_id = as.numeric(factor(paste(cntry, essround, sep = "_"), 
                                      levels = unique(paste(cntry, essround, sep = "_"))))
    )
```

## Year of the ESS round

```{r}
ess_filtered <- ess_filtered |>
  mutate(year = case_when(
    essround == 1 ~ 2002,
    essround == 2 ~ 2004,
    essround == 3 ~ 2006,
    essround == 4 ~ 2008,
    essround == 5 ~ 2010,
    essround == 6 ~ 2012,
    essround == 7 ~ 2014,
    essround == 8 ~ 2016,
    essround == 9 ~ 2018,
    essround == 10 ~ 2020,
    essround == 11 ~ 2023
  ))
```

## Left-right scale and "Don't know" dummy

```{r}
ess_filtered <- ess_filtered |>
  mutate(
    # Convert lrscale to numeric
    lrscale_numeric = as.numeric(lrscale),
    # Recode the left-right scale (NA for 77, 88, 99)
    leftright = case_when(
      lrscale_numeric %in% c(77, 88, 99) ~ NA_real_,
      TRUE ~ -lrscale_numeric + 10
    ),
    # "Don't know" dummy (NA for 77 and 99)
    lrscale_dontknow = case_when(
      lrscale == 88 ~ 1,
      lrscale %in% c(77, 99) ~ NA_real_,
      TRUE ~ 0
    )
  )
```

## Recode issue items

```{r}
ess_filtered <- ess_filtered |>
  mutate(
    # Redistribution
    redistribution = case_when(
      gincdif == 5 ~ 0,
      gincdif == 4 ~ 1,
      gincdif == 3 ~ 2,
      gincdif == 2 ~ 3,
      gincdif == 1 ~ 4,
      gincdif %in% c(7, 8, 9) ~ NA_real_
    ),
    
    # Homosexuals
    homosexuals = case_when(
      freehms == 5 ~ 0,
      freehms == 4 ~ 1,
      freehms == 3 ~ 2,
      freehms == 2 ~ 3,
      freehms == 1 ~ 4,
      freehms %in% c(7, 8, 9) ~ NA_real_
    ),
    
    # Environment
    environment = case_when(
      impenv == 6 ~ 0, # "not like me at all"
      impenv == 5 ~ 1,
      impenv == 4 ~ 2,
      impenv == 3 ~ 3,
      impenv == 2 ~ 4,
      impenv == 1 ~ 5, # "very much like me"
      impenv %in% c(7, 8, 9) ~ NA_real_
    ),
    
    # Immigration (reversed)
    immigration = case_when(
      impcntr == 4 ~ 0, # Allow none
      impcntr == 3 ~ 1, # Allow a few
      impcntr == 2 ~ 2, # Allow some
      impcntr == 1 ~ 3, # Allow many
      impcntr %in% c(7, 8, 9) ~ NA_real_ # Missing Values
    )
  )

# Prepare data for factor analysis
immigration_items <- ess_filtered |>
  mutate(
    imsmetn = ifelse(imsmetn %in% c(7, 8, 9), NA_real_, imsmetn),
    imdfetn = ifelse(imdfetn %in% c(7, 8, 9), NA_real_, imdfetn),
    imueclt = ifelse(imueclt %in% c(77, 88, 99), NA_real_, imueclt),
    imbgeco = ifelse(imbgeco %in% c(77, 88, 99), NA_real_, imbgeco),
    imwbcnt = ifelse(imwbcnt %in% c(77, 88, 99), NA_real_, imwbcnt)
  ) |>
  select(imsmetn, imdfetn, immigration, imueclt, imbgeco, imwbcnt)

# Check the distribution of missing values (optional)
sapply(immigration_items, function(x) sum(is.na(x)))

# Use only complete cases
valid_rows <- complete.cases(immigration_items)

# Factor analysis
factor_res <- factanal(immigration_items[valid_rows, ], factors = 1, scores = "regression")

# Extract and reverse factor scores
ess_filtered <- ess_filtered |>
  mutate(
    immigration_fact = ifelse(row_number() %in% which(valid_rows), -factor_res$scores[, 1], NA),
    immigration_fact_std = ifelse(
      !is.na(immigration_fact),
      round((immigration_fact - min(immigration_fact, na.rm = TRUE)) /
        (max(immigration_fact, na.rm = TRUE) - min(immigration_fact, na.rm = TRUE)), 6),
      NA
    )
  )

# Standardize other variables
ess_filtered <- ess_filtered |>
  mutate(
    redistribution_std = round(redistribution / 4, 6),
    homosexuals_std = round(homosexuals / 4, 6),
    environment_std = round(environment / 5, 6),
    immigration_std = round(immigration / 3, 6) # Verwende umgekehrte `immigration`
  )
```

## Define cohorts: Mitteregger (2024)

```{r}
# Define cohorts with missing value handling
ess_filtered <- ess_filtered |>
  mutate(
    yrbrn_clean = ifelse(yrbrn %in% c(6666, 7777, 8888, 9999), NA, yrbrn),
    cohort = case_when(
      yrbrn_clean <= 1925 ~ 0,
      yrbrn_clean >= 1926 & yrbrn_clean <= 1945 ~ 1,
      yrbrn_clean >= 1946 & yrbrn_clean <= 1957 ~ 2,
      yrbrn_clean >= 1958 & yrbrn_clean <= 1968 ~ 3,
      yrbrn_clean >= 1969 & yrbrn_clean <= 1979 ~ 4,
      yrbrn_clean >= 1980 ~ 5,
      TRUE ~ NA_real_ # Ensure unexpected values are NA
    )
  )

# Define cohort labels
cohort_labels <- c("interwar", "post-war", "60-70s", "80s", "90s", "millennial")

# Assign labels to cohorts
ess_filtered <- ess_filtered |>
  mutate(cohort = labelled(cohort, labels = setNames(0:5, cohort_labels)))
```

## Define cohorts: Decade of Birth

```{r}
ess_filtered <- ess_filtered |>
  mutate(
    yrbrn_clean = ifelse(yrbrn %in% c(6666, 7777, 8888, 9999), NA, yrbrn),
    cohort_alt = case_when(
      yrbrn_clean <= 1929 ~ 0,
      yrbrn_clean >= 1930 & yrbrn_clean <= 1939 ~ 1,
      yrbrn_clean >= 1940 & yrbrn_clean <= 1949 ~ 2,
      yrbrn_clean >= 1950 & yrbrn_clean <= 1959 ~ 3,
      yrbrn_clean >= 1960 & yrbrn_clean <= 1969 ~ 4,
      yrbrn_clean >= 1970 & yrbrn_clean <= 1979 ~ 5,
      yrbrn_clean >= 1980 & yrbrn_clean <= 1989 ~ 6,
      yrbrn_clean >= 1990 ~ 7,
      TRUE ~ NA_real_ # Ensure unexpected values are NA
    )
  )

# Define cohort labels
cohort_alt_labels <- c("-1929", "1930s", "1940s", "1950s", "1960s", "1970s", "1980s", "1990-")

# Assign labels to cohorts
ess_filtered <- ess_filtered |>
  mutate(cohort_alt = labelled(cohort_alt, labels = setNames(0:7, cohort_alt_labels)))
```

## Age groups

```{r}
ess_filtered <- ess_filtered |>
  mutate(
     # Clean age variable by removing all missing values
    agea_clean = ifelse(agea %in% c(999, 6666, 7777, 8888), NA_real_, agea),
    # Create age groups
    age_group = case_when(
      agea_clean > 17 & agea_clean <= 29 ~ 1,
      agea_clean >= 30 & agea_clean <= 65 ~ 2,
      agea_clean >= 66 ~ 3,
      TRUE ~ NA_real_ # Ensure unexpected values are NA
    )
  )

# Assign labels to age groups
age_group_labels <- c("18-29", "30-65", "66+")
ess_filtered <- ess_filtered |>
  mutate(age_group = labelled(age_group, labels = setNames(0:3, age_group_labels)))
```

## Recode education

```{r}
ess_filtered <- ess_filtered |>
  mutate(
    # Clean edulvla from missing values
    edulvla_clean = ifelse(edulvla %in% c(77, 88, 99), NA_real_, edulvla),
    # Clean edulvlb from missing values
    edulvlb_clean = ifelse(edulvlb %in% c(7777, 8888, 9999), NA_real_, edulvlb),
    
    # Recoding edulvla
    education_a = recode(edulvla_clean, 
                         `1` = 0,  # Low education
                         `2` = 0,  # Low education
                         `3` = 1,  # Medium education
                         `4` = 1,  # Medium education
                         `5` = 2,  # High education
                         `55` = NA_real_),  # Other
    
    # Recoding edulvlb
    education_b = case_when(
      edulvlb_clean >= 0 & edulvlb_clean <= 229 ~ 0,  # Low education
      edulvlb_clean >= 300 & edulvlb_clean <= 423 ~ 1,  # Medium education
      edulvlb_clean >= 500 & edulvlb_clean <= 800 ~ 2,  # High education
      edulvlb_clean == 5555 ~ NA_real_,  # Other
      TRUE ~ NA_real_  # Fallback
    ),
    
    # Combined education variable
    education = ifelse(is.na(education_a), education_b, education_a),
    edulowhigh = case_match(eisced,
                            c(0:4)~"low",
                            c(5:7)~"high")
  )
```

## Immigration background

```{r}
ess_filtered <- ess_filtered |>
  mutate(
    # Clean variables from missing values
    brncntr_clean = ifelse(brncntr %in% c(7, 8, 9), NA_real_, brncntr),
    facntr_clean = ifelse(facntr %in% c(7, 8, 9), NA_real_, facntr),
    mocntr_clean = ifelse(mocntr %in% c(7, 8, 9), NA_real_, mocntr),
    
    # Calculate immigration background
    immigrant = case_when(
      brncntr_clean == 1 & facntr_clean == 1 & mocntr_clean == 1 ~ 0, # No immigration background
      brncntr_clean == 2 | facntr_clean == 2 | mocntr_clean == 2 ~ 1, # Immigration background
      TRUE ~ NA_real_ # Ensure unexpected cases are set to NA
    )
  )
```

## Recode voting decision and country-specific recodes

```{r}
# Recode voting decision
ess_filtered <- ess_filtered |>
  mutate(voted = NA_integer_) # Initialize variable

# Define labels
voted_labels <- c(
  "Abstention", "Communist/Socialist", "Social democracy", "Green",
  "Liberal", "Conservative/Christian democracy", "Radical Right",
  "Agrarian", "Special issue", "Other", "Blank / Invalid",
  "Not eligible to vote"
)
voted <- labelled(NA_integer_, labels = setNames(0:11, voted_labels))

# Recode based on the "vote" variable
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      vote == 2 ~ 0,  # Abstention
      vote == 3 ~ 11, # Not eligible to vote
      TRUE ~ voted    # Keep original value
    )
  )

# Country-specific recodes
# Belgium
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Communist/Socialist
      prtvtbe %in% c(7, 16) | 
        prtvtcbe %in% c(6, 14) | 
        prtvtdbe %in% c(6, 14) | 
        prtvtebe %in% c(5, 12) ~ 1, 
      
      # Social democracy
      prtvtbe %in% c(3, 14) | 
        prtvtabe %in% c(5, 13) | 
        prtvtbbe %in% c(5, 13) | 
        prtvtcbe %in% c(5, 13) | 
        prtvtdbe %in% c(5, 13) | 
        prtvtebe %in% c(4, 11) ~ 2, 
      
      # Green
      prtvtbe %in% c(1, 11) | 
        prtvtabe %in% c(1, 10) | 
        prtvtbbe %in% c(1, 10) | 
        prtvtcbe %in% c(1, 10) | 
        prtvtdbe %in% c(1, 10) | 
        prtvtebe %in% c(1, 9) ~ 3, 
      
      # Liberal
      prtvtbe %in% c(5, 9, 13) | 
        prtvtabe %in% c(6, 8, 12) | 
        prtvtbbe %in% c(4, 8, 12) | 
        prtvtcbe %in% c(4, 8, 12) | 
        prtvtdbe %in% c(4, 8, 12, 16) | 
        prtvtebe %in% c(10, 7, 13) ~ 4, 
      
      # Conservative/Christian democracy
      prtvtbe %in% c(2, 12) | 
        prtvtabe %in% c(2, 3, 9) | 
        prtvtbbe %in% c(2, 9) | 
        prtvtcbe %in% c(2, 3, 9, 15) | 
        prtvtdbe %in% c(2, 3, 9, 15) | 
        prtvtebe %in% c(2, 3, 8) ~ 5, 
      
      # Radical Right
      prtvtbe %in% c(6, 8, 15) | 
        prtvtabe %in% c(7, 11) | 
        prtvtbbe %in% c(7, 11) | 
        prtvtcbe %in% c(7, 11) | 
        prtvtdbe %in% c(7, 11) | 
        prtvtebe %in% c(6) ~ 6, 
      
      # Other
      prtvtbe %in% c(17) | 
        prtvtabe %in% c(14) | 
        prtvtbbe %in% c(14) | 
        prtvtcbe %in% c(16) ~ 9, 
      
      # Blank / Invalid
      prtvtabe %in% c(15, 16) | 
        prtvtbbe %in% c(15, 16) | 
        prtvtcbe %in% c(17, 18) |
        prtvtebe %in% c(14, 15) ~ 10,
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )

# Switzerland: Recode with additional waves 8, 9, and 10
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Communist/Socialist
      prtvtbch %in% c(7) | 
        prtvtach %in% c(9) | 
        prtvtch %in% c(9) | 
        prtvtcch %in% c(7) | 
        prtvtdch %in% c(11, 14) | 
        prtvtech %in% c(11, 14) | 
        prtvtfch %in% c(11) | 
        prtvthch %in% c(11) ~ 1, 
      
      # Social democracy
      prtvtch %in% c(3, 6) | 
        prtvtach %in% c(3) | 
        prtvtbch %in% c(3) | 
        prtvtcch %in% c(3) | 
        prtvtdch %in% c(2) | 
        prtvtech %in% c(2) | 
        prtvtfch %in% c(2) | 
        prtvtgch %in% c(2) | 
        prtvthch %in% c(2) ~ 2, 
      
      # Green
      prtvtch %in% c(10, 14) | 
        prtvtach %in% c(10, 14) | 
        prtvtbch %in% c(8, 9) | 
        prtvtcch %in% c(8, 9) | 
        prtvtdch %in% c(5, 6) | 
        prtvtech %in% c(5, 6, 15) | 
        prtvtfch %in% c(5, 6) | 
        prtvtgch %in% c(5, 6, 13) | 
        prtvthch %in% c(4, 6, 12) ~ 3, 
      
      # Liberal
      prtvtch %in% c(1, 5) | 
        prtvtach %in% c(1, 5) | 
        prtvtbch %in% c(1, 5) | 
        prtvtcch %in% c(1, 5) | 
        prtvtdch %in% c(3, 16) | 
        prtvtech %in% c(3) | 
        prtvtfch %in% c(3) | 
        prtvtgch %in% c(3) | 
        prtvthch %in% c(3) ~ 4, 
      
      # Conservative/Christian democracy
      prtvtch %in% c(2, 7, 8, 12) | 
        prtvtach %in% c(2, 7, 8, 12) | 
        prtvtbch %in% c(2, 6, 11, 12) | 
        prtvtcch %in% c(2, 6, 11, 12) | 
        prtvtdch %in% c(4, 7, 8, 9, 13) | 
        prtvtech %in% c(4, 7, 8, 9, 13) | 
        prtvtfch %in% c(4, 7, 8, 9) | 
        prtvtgch %in% c(4, 7, 8, 9) | 
        prtvthch %in% c(5, 7, 8, 9, 13) ~ 5, 
      
      # Radical Right
      prtvtch %in% c(4, 11, 13, 15) | 
        prtvtach %in% c(4, 11, 13, 15) | 
        prtvtbch %in% c(4, 10, 13) | 
        prtvtcch %in% c(4, 10, 13) | 
        prtvtdch %in% c(1, 10, 12, 15, 17) | 
        prtvtech %in% c(1, 10, 12) | 
        prtvtfch %in% c(10, 1) | 
        prtvtgch %in% c(10, 1) | 
        prtvthch %in% c(10, 1, 15) ~ 6, 
      
      # Other
      prtvtch %in% c(16) | 
        prtvtach %in% c(16) | 
        prtvtbch %in% c(14) | 
        prtvtcch %in% c(20) | 
        prtvtech %in% c(18) | 
        prtvtfch %in% c(16, 18) | 
        prtvtgch %in% c(15) | 
        prtvthch %in% c(17) ~ 9, 
      
      # Blank / Invalid
      prtvtbch %in% c(15, 16) | 
        prtvtcch %in% c(18, 19) | 
        prtvtdch %in% c(19, 20) | 
        prtvtech %in% c(19, 20) | 
        prtvtfch %in% c(19, 55) | 
        prtvtgch %in% c(16) | 
        prtvthch %in% c(18, 20) ~ 10, 
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )

# Germany: Recode including new variable prtvgde2 (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Party of Democratic Socialism (PDS)
        prtvde2  %in% c(5) |
        prtvade2 %in% c(5) |
        prtvbde2 %in% c(5) | 
        prtvcde2 %in% c(5) | 
        prtvdde2 %in% c(5) | 
        prtvede2 %in% c(3) | 
        prtvfde2 %in% c(3) | 
        prtvgde2 %in% c(3) ~ 1, 
      
      # Social Democratic Party (SPD)
      prtvde2 %in% c(1) |
        prtvade2 %in% c(1) |
        prtvbde2 %in% c(1) | 
        prtvcde2 %in% c(1) |
        prtvdde2 %in% c(1) |
        prtvede2 %in% c(2) | 
        prtvfde2 %in% c(2) | 
        prtvgde2 %in% c(2) ~ 2, 
      
      # Alliance 90/The Greens
      prtvde2 %in% c(3) |
        prtvade2 %in% c(3) |
        prtvbde2 %in% c(3) | 
        prtvcde2 %in% c(3) |
        prtvdde2 %in% c(3) |
        prtvede2 %in% c(4) | 
        prtvfde2 %in% c(4) | 
        prtvgde2 %in% c(4) ~ 3, 
      
      # Free Democratic Party (FDP)
      prtvde2 %in% c(4) |
        prtvade2 %in% c(4) |
        prtvbde2 %in% c(4) | 
        prtvcde2 %in% c(4) |
        prtvdde2 %in% c(4, 8) |
        prtvede2 %in% c(5, 7) | 
        prtvfde2 %in% c(5) | 
        prtvgde2 %in% c(5) ~ 4, 
      
      # Christian Democratic Union/Christian Social Union (CDU/CSU)
      prtvde2 %in% c(2) |
        prtvade2 %in% c(2) |
        prtvbde2 %in% c(2) | 
        prtvcde2 %in% c(2) |
        prtvdde2 %in% c(2) |
        prtvede2 %in% c(1) | 
        prtvfde2 %in% c(1) | 
        prtvgde2 %in% c(1, 7) ~ 5, # Freie Wähler/Free Voters (7) 
      
      # The Republicans (REP)
      prtvde2 %in% c(6) |
        prtvade2 %in% c(6, 7) |
        prtvbde2 %in% c(6, 7) | 
        prtvcde2 %in% c(6, 7) |
        prtvdde2 %in% c(6, 7) |
        prtvede2 %in% c(6, 8) | 
        prtvfde2 %in% c(6) | 
        prtvgde2 %in% c(6) ~ 6, # AfD (6)
      
      # Other
      prtvde2 %in% c(7) |
        prtvade2 %in% c(8) |
        prtvbde2 %in% c(8) | 
        prtvcde2 %in% c(8) |
        prtvdde2 %in% c(9) |
        prtvede2 %in% c(9) | 
        prtvfde2 %in% c(7) | 
        prtvgde2 %in% c(8, 9, 55) ~ 9,  # dieBasis (8), Die PARTEI (9), Other (55) 
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )

# Norway: Recode including new variable prtvtcno (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Rødt / Sosialistisk Venstreparti (SV)
      prtvtno %in% c(1, 2) | 
        prtvtano %in% c(1, 2) | 
        prtvtbno %in% c(1, 2) | 
        prtvtcno %in% c(1, 2) ~ 1,  # Rødt (1), SV (2)
      
      # Det norske Arbeiderparti (Labour Party)
      prtvtno %in% c(3) | 
        prtvtano %in% c(3) | 
        prtvtbno %in% c(3) | 
        prtvtcno %in% c(3) ~ 2,  # Arbeiderpartiet (3)
      
      # Miljøpartiet De Grønne (Green Party)
      prtvtbno %in% c(10) | 
        prtvtcno %in% c(9) ~ 3,  # MDG (9)
      
      # Venstre (Liberal Party)
      prtvtno %in% c(4) | 
        prtvtano %in% c(4) | 
        prtvtbno %in% c(4) | 
        prtvtcno %in% c(4) ~ 4,  # Venstre (4)
      
      # Kristelig Folkeparti (KrF) / Høyre (Conservatives)
      prtvtno %in% c(5, 7, 9) | 
        prtvtano %in% c(5, 7, 9) | 
        prtvtbno %in% c(5, 7) | 
        prtvtcno %in% c(5, 7) ~ 5,  # KrF (5), Høyre (7)
      
      # Fremskrittspartiet (Progress Party)
      prtvtno %in% c(8) | 
        prtvtano %in% c(8) | 
        prtvtbno %in% c(8) | 
        prtvtcno %in% c(8) ~ 6,  # Fremskrittspartiet (8)
      
      # Senterpartiet (Centre Party)
      prtvtno %in% c(6) | 
        prtvtano %in% c(6) | 
        prtvtbno %in% c(6, 9) | 
        prtvtcno %in% c(6) ~ 7,  # Senterpartiet (6)
      
      # Other (inkl. Pasientfokus)
      prtvtno %in% c(10) | 
        prtvtano %in% c(10) | 
        prtvtbno %in% c(11) | 
        prtvtcno %in% c(10, 11) ~ 9,  # Pasientfokus (10), Other (11)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# United Kingdom
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Sinn Féin (nir)
      prtvtgb %in% c(13) |
        prtvtagb %in% c(13) |
        prtvtbgb %in% c(11) | 
        prtvtcgb %in% c(11) |
        prtvtdgb %in% c(12) ~ 1,
      
      # Labour / Scottish National Party / Social Democratic and Labour Party (nir)
      prtvtgb %in% c(2, 4, 14) |
        prtvtagb %in% c(2, 4, 14) |
        prtvtbgb %in% c(2, 4, 12) | 
        prtvtcgb %in% c(2, 4, 12) |
        prtvtdgb %in% c(2, 4) ~ 2,
      
      # Green Party
      prtvtgb %in% c(6) |
        prtvtagb %in% c(6) |
        prtvtbgb %in% c(6) | 
        prtvtcgb %in% c(6, 15) |
        prtvtdgb %in% c(6, 15) ~ 3,
      
      # Liberal Democrat / Alliance Party (nir)
      prtvtgb %in% c(3, 15) |
        prtvtagb %in% c(3, 15) |
        prtvtbgb %in% c(3, 13) | 
        prtvtcgb %in% c(3, 13) |
        prtvtdgb %in% c(3) ~ 4,
      
      # Conservative / Ulster Unionist Party / Democratic Unionist Party (nir)
      prtvtgb %in% c(1, 11, 12, 17) |
        prtvtagb %in% c(1, 11, 12) |
        prtvtbgb %in% c(1, 9, 10) | 
        prtvtcgb %in% c(1, 9, 10) |
        prtvtdgb %in% c(1, 10, 11) ~ 5,
      
      # British National Party / UK Independence Party / Brexit Party
      prtvtagb %in% c(7, 8) |
        prtvtbgb %in% c(7) |
        prtvtcgb %in% c(7) | 
        prtvtdgb %in% c(7, 8) ~ 6,
      
      # Plaid Cymru
      prtvtgb %in% c(5) |
        prtvtagb %in% c(5) |
        prtvtbgb %in% c(5) | 
        prtvtcgb %in% c(5) |
        prtvtdgb %in% c(5) ~ 8,
      
      # Other
      prtvtgb %in% c(7, 18, 22) |
        prtvtagb %in% c(22) |
        prtvtbgb %in% c(8, 14, 16, 18) |
        prtvtcgb %in% c(16, 19) |
        prtvtdgb %in% c(9, 23) ~ 9,
      
       # Fallback: Keep existing values
      TRUE ~ voted
    )
  )

# Sweden
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Left
      prtvtse %in% c(7) |
        prtvtase %in% c(7) |
        prtvtbse %in% c(7) | 
        prtvtcse %in% c(7) |
        prtvtdse %in% c(8) ~ 1,
      
      # Social Democrats
      prtvtse %in% c(6) |
        prtvtase %in% c(6) |
        prtvtbse %in% c(6) | 
        prtvtcse %in% c(6) |
        prtvtdse %in% c(6) ~ 2,
      
      # Green Party
      prtvtse %in% c(4) |
        prtvtase %in% c(4) |
        prtvtbse %in% c(4) | 
        prtvtcse %in% c(4) |
        prtvtdse %in% c(4) ~ 3,
      
      # Liberals
      prtvtse %in% c(2) |
        prtvtase %in% c(2) |
        prtvtbse %in% c(2) | 
        prtvtcse %in% c(2) |
        prtvtdse %in% c(3) ~ 4,
      
      # Christian Democrats / Conservative
      prtvtse %in% c(3, 5) |
        prtvtase %in% c(3, 5, 9) |
        prtvtbse %in% c(3, 5) | 
        prtvtcse %in% c(3, 5) |
        prtvtdse %in% c(2, 5) ~ 5,
      
      # Sverigedemokraterna
      prtvtase %in% c(10) |
        prtvtbse %in% c(10) | 
        prtvtcse %in% c(9) |
        prtvtdse %in% c(7) ~ 6,
      
      # Centre Party
      prtvtse %in% c(1) |
        prtvtase %in% c(1) |
        prtvtbse %in% c(1) | 
        prtvtcse %in% c(1) |
        prtvtdse %in% c(1) ~ 7,
      
      # FI (Feministiskt initiativ) / Piratpartiet
      prtvtase %in% c(8) |
        prtvtbse %in% c(8, 9) | 
        prtvtcse %in% c(8) ~ 8,
      
      # Other
      prtvtse %in% c(8) |
        prtvtase %in% c(11) |
        prtvtbse %in% c(11) | 
        prtvtcse %in% c(10) |
        prtvtdse %in% c(9) ~ 9,
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )

# France: Recode including new variable prtvtffr (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Lutte Ouvrière (LO) / Nouveau Parti Anti-Capitaliste (NPA) / Parti Communiste / NUPES
      prtvtfr %in% c(4, 5, 6, 9) |
        prtvtafr %in% c(4, 5, 6, 9) |
        prtvtbfr %in% c(3, 4, 7, 14) |
        prtvtcfr %in% c(4, 5, 6) | 
        prtvtdfr %in% c(1, 2, 3, 4) |
        prtvtefr %in% c(1, 2, 3, 4) |
        prtvtffr %in% c(1, 2, 3) ~ 1,  # LO (1), NPA (2), NUPES (3) -> coded as Communist/Socialist because of LFI being the largest party, but: ATTENTION, also includes parties of social-democratic and green party family (better exclude France ESS wave 11 when analyzing voting for any of the three involved parties)
      
      # Parti Socialiste (PS)
      prtvtfr %in% c(10) |
        prtvtafr %in% c(10) | 
        prtvtbfr %in% c(8, 9) |
        prtvtcfr %in% c(7, 9) | 
        prtvtdfr %in% c(5) |
        prtvtefr %in% c(5) ~ 2,  
      
      # Les Verts / EELV
      prtvtfr %in% c(14, 15) |
        prtvtafr %in% c(14, 15) | 
        prtvtbfr %in% c(12, 13) |
        prtvtcfr %in% c(12, 13) | 
        prtvtdfr %in% c(6) |
        prtvtefr %in% c(6) ~ 3,
      
      # Ensemble! (LREM, MODEM, Horizons) (ehemals LREM)
      prtvtbfr %in% c(6) |
        prtvtcfr %in% c(1) | 
        prtvtdfr %in% c(7) |
        prtvtefr %in% c(7) |
        prtvtffr %in% c(5) ~ 4,  # Ensemble! (5)
      
      # Les Républicains, UDI, Debout la France
      prtvtfr %in% c(2, 8, 11, 12, 13, 1) |
        prtvtafr %in% c(2, 8, 11, 12, 13, 1) | 
        prtvtbfr %in% c(1, 5, 10, 11) |
        prtvtcfr %in% c(3, 8, 10, 11) | 
        prtvtdfr %in% c(8, 9, 10) |
        prtvtefr %in% c(8, 9, 10) |
        prtvtffr %in% c(6, 7) ~ 5,  # Les Républicains & UDI (6), Union pour la France (7)
      
      # Rassemblement National (ex-Front National) / Reconquête!
      prtvtfr %in% c(3, 7) |
        prtvtafr %in% c(3, 7) | 
        prtvtbfr %in% c(2, 15) |
        prtvtcfr %in% c(2) | 
        prtvtdfr %in% c(11) |
        prtvtefr %in% c(11) |
        prtvtffr %in% c(8, 9) ~ 6,  # Rassemblement National (8), Reconquête! (9)
      
      # Other
      prtvtfr %in% c(16) |
        prtvtafr %in% c(16) | 
        prtvtbfr %in% c(16) |
        prtvtcfr %in% c(14) | 
        prtvtdfr %in% c(12) |
        prtvtefr %in% c(12) |
        prtvtffr %in% c(31) ~ 9,  # Other (31)
      
      # Blank, Nul
      prtvtafr %in% c(17) |
        prtvtbfr %in% c(17, 18) | 
        prtvtcfr %in% c(15, 16) |
        prtvtffr %in% c(10, 11) ~ 10,  # Blanc (10), Nul (11)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Netherlands: Recode including new variable prtvtinl (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Socialist Party
      prtvtnl %in% c(7) |
        prtvtanl %in% c(7) |
        prtvtbnl %in% c(7) | 
        prtvtcnl %in% c(7) |
        prtvtdnl %in% c(5) |
        prtvtenl %in% c(5) | 
        prtvtfnl %in% c(4) |
        prtvtgnl %in% c(4) |
        prtvthnl %in% c(4) |
        prtvtinl %in% c(4) ~ 1,  # Socialist Party (4)
      
      # Labour Party (PvdA)
      prtvtnl %in% c(2) |
        prtvtanl %in% c(2) |
        prtvtbnl %in% c(2) | 
        prtvtcnl %in% c(2) |
        prtvtdnl %in% c(2) |
        prtvtenl %in% c(2) | 
        prtvtfnl %in% c(2) |
        prtvtgnl %in% c(2) |
        prtvthnl %in% c(2) |
        prtvtinl %in% c(2) ~ 2,  # Labour Party (2)
      
      # Green Left (GroenLinks)
      prtvtnl %in% c(6) |
        prtvtanl %in% c(6) |
        prtvtbnl %in% c(6) | 
        prtvtcnl %in% c(6) |
        prtvtdnl %in% c(7) |
        prtvtenl %in% c(7) | 
        prtvtfnl %in% c(8) |
        prtvtgnl %in% c(8) |
        prtvthnl %in% c(8) |
        prtvtinl %in% c(8) ~ 3,  # Green Left (8)
      
      # People's Party for Freedom and Democracy (VVD), Democrats '66 (D66)
      prtvtnl %in% c(3, 5) |
        prtvtanl %in% c(3, 5) |
        prtvtbnl %in% c(3, 5) | 
        prtvtcnl %in% c(3, 5) |
        prtvtdnl %in% c(1, 6, 11) |
        prtvtenl %in% c(1, 6, 11, 12) | 
        prtvtfnl %in% c(1, 6, 11) |
        prtvtgnl %in% c(1, 6, 11, 12) |
        prtvthnl %in% c(1, 6, 11, 12) |
        prtvtinl %in% c(1, 6) ~ 4,  # VVD (1), D66 (6)
      
      # Christian Democratic Appeal (CDA), Christian Union (CU)
      prtvtnl %in% c(1, 8, 10) |
        prtvtanl %in% c(1, 8, 10) |
        prtvtbnl %in% c(1, 8, 10) | 
        prtvtcnl %in% c(1, 8, 10) |
        prtvtdnl %in% c(4, 8, 9) |
        prtvtenl %in% c(4, 8, 9) | 
        prtvtfnl %in% c(5, 7, 9) |
        prtvtgnl %in% c(5, 7, 9) |
        prtvthnl %in% c(5, 7, 9) |
        prtvtinl %in% c(5, 7, 9) ~ 5,  # CDA (5), CU (7), SGP (9)
      
      # Party for Freedom (PVV), Forum for Democracy (FvD), JA21
      prtvtnl %in% c(4, 9) |
        prtvtanl %in% c(4, 9) |
        prtvtbnl %in% c(4, 9) | 
        prtvtcnl %in% c(4, 9, 11) |
        prtvtdnl %in% c(3) |
        prtvtenl %in% c(3) | 
        prtvtfnl %in% c(3) |
        prtvtgnl %in% c(3, 13) |
        prtvthnl %in% c(3, 13, 16) |
        prtvtinl %in% c(3, 13, 16) ~ 6,  # PVV (3), FvD (13), JA21 (16)
      
      # BoerBurgerBeweging (BBB)
      prtvthnl %in% c(17) |
        prtvtinl %in% c(17) ~ 7,  # BBB (17)
      
      # Party for the Animals (PvdD), Volt, 50PLUS
      prtvtanl %in% c(11) |
        prtvtcnl %in% c(12) |
        prtvtdnl %in% c(10) | 
        prtvtenl %in% c(10) |
        prtvtfnl %in% c(10) |
        prtvtgnl %in% c(10) | 
        prtvthnl %in% c(10, 15) |
        prtvtinl %in% c(10, 11, 15) ~ 8,  # PvdD (10), 50PLUS (11), Volt (15)
      
      # Other
      prtvtnl %in% c(11) |
        prtvtanl %in% c(12) |
        prtvtbnl %in% c(11) | 
        prtvtcnl %in% c(13) |
        prtvtdnl %in% c(12) |
        prtvtenl %in% c(13) | 
        prtvtfnl %in% c(16) |
        prtvtgnl %in% c(14, 16) |
        prtvthnl %in% c(14, 31) |
        prtvtinl %in% c(12, 14, 31) ~ 9,  # DENK (12), Artikel1/Bij1 (14), Other (31)
      
      # Blanc
      prtvtanl %in% c(13) |
        prtvtbnl %in% c(12) |
        prtvtcnl %in% c(14) | 
        prtvtdnl %in% c(13) |
        prtvtenl %in% c(14) |
        prtvtfnl %in% c(17) | 
        prtvtgnl %in% c(17) |
        prtvthnl %in% c(18) |
        prtvtinl %in% c(18) ~ 10,  # Blanc (18)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Ireland: Recode including new variable prtvteie (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Sinn Féin, Socialist Party, Solidarity/PBP, Independent
      prtvtie %in% c(6) |
        prtvtaie %in% c(6, 7, 8, 9) | 
        prtvtbie %in% c(1, 7, 9) |
        prtvtcie %in% c(1, 7, 9) | 
        prtvtdie %in% c(4, 7, 8, 11, 13) |
        prtvteie %in% c(4, 7, 8) ~ 1,  # Sinn Féin (4), Socialist Party (7), Solidarity/PBP (8)
      
      # Labour, Social Democrats
      prtvtie %in% c(3) |
        prtvtaie %in% c(5) | 
        prtvtbie %in% c(6, 8) |
        prtvtcie %in% c(6, 8) | 
        prtvtdie %in% c(3, 12) |
        prtvteie %in% c(3, 11) ~ 2,  # Labour (3), Social Democrats (11)
      
      # Green Party
      prtvtie %in% c(5) |
        prtvtaie %in% c(3) | 
        prtvtbie %in% c(4) |
        prtvtcie %in% c(4) | 
        prtvtdie %in% c(6) |
        prtvteie %in% c(6) ~ 3,  # Green Party (6)
      
      # Progressive Democrats (Nicht mehr existent, entfernt)
      
      # Fianna Fáil, Fine Gael, Renua Ireland, Aontu
      prtvtie %in% c(1, 2) |
        prtvtaie %in% c(1, 2) | 
        prtvtbie %in% c(2, 3) |
        prtvtcie %in% c(2, 3) | 
        prtvtdie %in% c(1, 2, 9, 10) |
        prtvteie %in% c(1, 2, 9, 10) ~ 5,  # Fianna Fáil (1), Fine Gael (2), Renua (9), Aontu (10)
      
      # Other
      prtvtie %in% c(8) |
        prtvtaie %in% c(10) | 
        prtvtbie %in% c(5, 10) |
        prtvtcie %in% c(5, 10) | 
        prtvtdie %in% c(5, 14) |
        prtvteie %in% c(5, 12) ~ 9,  # , Independent (5), Other (12)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Spain: Recode including new variable prtvtges (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Unidas Podemos (eingegangen in SUMAR), Izquierda Unida, Compromís, Regional alliances
      prtvtes %in% c(3) |
        prtvtaes %in% c(3) |
        prtvtbes %in% c(3) | 
        prtvtces %in% c(4, 5, 11) |
        prtvtdes %in% c(3, 7, 8, 12) | 
        prtvtees %in% c(3, 7, 9, 19) |
        prtvtfes %in% c(4, 5, 10) |
        prtvtges %in% c(4) ~ 1,  # SUMAR (4)
      
      # PSOE
      prtvtes %in% c(2, 11) |
        prtvtaes %in% c(2, 11) |
        prtvtbes %in% c(2) | 
        prtvtces %in% c(2) |
        prtvtdes %in% c(2) |
        prtvtees %in% c(2) | 
        prtvtfes %in% c(1) |
        prtvtges %in% c(2) ~ 2,  # PSOE (2)
      
      # ICV, Más País (aufgegangen in SUMAR)
      prtvtes %in% c(6) |
        prtvtaes %in% c(6) |
        prtvtees %in% c(8) | 
        prtvtfes %in% c(7, 15) |
        prtvtges %in% c(4) ~ 3,  # SUMAR (4)
      
      # UPyD, Ciudadanos (stark zurückgegangen, nicht mehr in Welle 11)
      
      # PP, PNV, CiU, Regional alliances
      prtvtes %in% c(1, 4, 7) |
        prtvtaes %in% c(1, 4, 7) |
        prtvtbes %in% c(1, 4, 6) | 
        prtvtces %in% c(1, 3, 7, 12) |
        prtvtdes %in% c(1, 11) |
        prtvtees %in% c(1, 11, 13, 17) | 
        prtvtfes %in% c(2, 9, 11, 14) |
        prtvtges %in% c(1, 8) ~ 5,  # PP (1), EAJ-PNV (8)
      
      # VOX
      prtvtees %in% c(16) |
        prtvtfes %in% c(3) |
        prtvtges %in% c(3) ~ 6,  # VOX (3)
      
      # ERC, Regional alliances
      prtvtes %in% c(5, 8, 9, 10, 12) |
        prtvtaes %in% c(5, 8, 9, 10, 12, 13) | 
        prtvtbes %in% c(7, 8, 9) |
        prtvtces %in% c(8, 9, 10) |
        prtvtdes %in% c(10, 14, 15, 18) | 
        prtvtees %in% c(10, 12, 14, 15, 18) |
        prtvtfes %in% c(8, 12, 13, 16, 18) |
        prtvtges %in% c(5, 6, 7, 9) ~ 8,  # ERC (5), Junts x Cat (6), EH-Bildu (7), BNG (9)
      
      # Other
      prtvtes %in% c(68) |
        prtvtaes %in% c(74) |
        prtvtbes %in% c(74) | 
        prtvtces %in% c(14) |
        prtvtdes %in% c(50) |
        prtvtees %in% c(20, 53) | 
        prtvtfes %in% c(17, 53) |
        prtvtges %in% c(12, 50) ~ 9,  # PACMA (12), Other (50)
      
      # Blank/Invalid
      prtvtaes %in% c(75, 76) |
        prtvtbes %in% c(75, 76) |
        prtvtces %in% c(15, 16) | 
        prtvtdes %in% c(51, 52) |
        prtvtees %in% c(54, 55) |
        prtvtfes %in% c(54, 55) |
        prtvtges %in% c(51, 52) ~ 10,  # Blank (51), Invalid (52)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Finland: Recode including new variable prtvtffi (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Left Alliance, Communist Party
      prtvtfi %in% c(7, 10, 11, 12) |
        prtvtafi %in% c(9, 10, 11) | 
        prtvtbfi %in% c(15, 16, 17, 18) |
        prtvtcfi %in% c(14, 15, 16) | 
        prtvtdfi %in% c(12, 13, 14) |
        prtvtefi %in% c(18, 19) |
        prtvtffi %in% c(7, 18) ~ 1,  # Left Alliance (7), Communist Party (18)
      
      # Social Democrats
      prtvtfi %in% c(9) |
        prtvtafi %in% c(8) |
        prtvtbfi %in% c(14, 11) | 
        prtvtcfi %in% c(13, 17) |
        prtvtdfi %in% c(11, 15) |
        prtvtefi %in% c(17) |
        prtvtffi %in% c(1) ~ 2,  # Social Democratic Party (1)
      
      # Green League
      prtvtfi %in% c(8) |
        prtvtafi %in% c(7) |
        prtvtbfi %in% c(13) | 
        prtvtcfi %in% c(12) |
        prtvtdfi %in% c(10) |
        prtvtefi %in% c(16) |
        prtvtffi %in% c(6) ~ 3,  # Green League (6)
      
      # Swedish People's Party, Liberals
      prtvtfi %in% c(2, 3) |
        prtvtbfi %in% c(2, 3) |
        prtvtcfi %in% c(2) | 
        prtvtdfi %in% c(2) |
        prtvtefi %in% c(2, 8) |
        prtvtffi %in% c(4, 9, 16) ~ 4,  # Swedish People's Party (4), The Liberal Party - Freedom to Choose, Movement Now (16)
      
      # National Coalition Party, Christian Democrats
      prtvtfi %in% c(1, 6) |
        prtvtafi %in% c(1, 2, 3, 6) |
        prtvtbfi %in% c(1, 6) | 
        prtvtcfi %in% c(1, 5) |
        prtvtdfi %in% c(1, 5) |
        prtvtefi %in% c(1, 7) |
        prtvtffi %in% c(3, 5) ~ 5,  # National Coalition Party (3), Christian Democrats (5)
      
      # True Finns (Perussuomalaiset)
      prtvtfi %in% c(5) |
        prtvtafi %in% c(5) |
        prtvtbfi %in% c(5, 8) | 
        prtvtcfi %in% c(4) |
        prtvtdfi %in% c(4) |
        prtvtefi %in% c(5) |
        prtvtffi %in% c(8) ~ 6,  # True Finns (8)
      
      # Centre Party
      prtvtfi %in% c(4) |
        prtvtafi %in% c(4) |
        prtvtbfi %in% c(4) | 
        prtvtcfi %in% c(3) |
        prtvtdfi %in% c(3) |
        prtvtefi %in% c(3) |
        prtvtffi %in% c(2) ~ 7,  # Centre Party (2)
      
      # Pirate Party, Independence Party
      prtvtcfi %in% c(8, 10) |
        prtvtdfi %in% c(7, 8) |
        prtvtefi %in% c(10) |
        prtvtffi %in% c(10) ~ 8,  # Pirate Party (10)
      
      # Other
      prtvtfi %in% c(13, 14) |
        prtvtafi %in% c(12) |
        prtvtbfi %in% c(9, 19) | 
        prtvtcfi %in% c(6, 7, 11, 18) |
        prtvtdfi %in% c(6, 9, 16) |
        prtvtefi %in% c(22) |
        prtvtffi %in% c(11, 12, 13, 14, 15, 17, 19, 20, 21, 22, 23, 24) ~ 9,  # Diverse kleinere Parteien
        
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Portugal: Recode including new variable prtvtept (Wave 11)
ess_filtered <- ess_filtered |>
  mutate(
    voted = case_when(
      # Bloco de Esquerda (BE), Kommunisten (PCP-PEV, PCTP/MRPP)
      prtvtpt %in% c(1, 5, 6, 8) |
        prtvtapt %in% c(1, 3, 4, 9) | 
        prtvtbpt %in% c(1, 3, 4, 9) |
        prtvtcpt %in% c(2, 3, 8) | 
        prtvtdpt %in% c(2, 3, 12) |
        prtvtept %in% c(3, 14, 15) ~ 1,  # Bloco de Esquerda (3), CDU (PCP-PEV) (14), PCTP/MRPP (15)
      
      # Partido Socialista (PS)
      prtvtpt %in% c(4, 10) |
        prtvtapt %in% c(6, 11) |
        prtvtbpt %in% c(6, 11) | 
        prtvtcpt %in% c(13) |
        prtvtdpt %in% c(7, 18) |
        prtvtept %in% c(9, 18) ~ 2,  # LIVRE(9), PS (18)
      
      # Pessoas-Animais-Natureza (PAN), LIVRE
      prtvtpt == 3 |
        prtvtcpt %in% c(5, 9, 15) |
        prtvtdpt %in% c(5, 11) |
        prtvtept %in% c(11, 13) ~ 3,  # MPT (11), PAN (13)
      
      # Partido Social Democrata (PSD), Alianca, Iniciativa Liberal
      prtvtpt == 11 |
        prtvtapt == 10 |
        prtvtbpt == 10 |
        prtvtcpt %in% c(16) | 
        prtvtdpt %in% c(16) |
        prtvtept %in% c(1, 7, 16) ~ 4,  # Iniciativa Liberal (IL) (7), PSD (16)
      
      # CDS-Partido Popular (CDS-PP), PPM
      prtvtpt %in% c(2, 9) |
        prtvtapt %in% c(2, 7) |
        prtvtbpt %in% c(2, 7) | 
        prtvtcpt %in% c(7, 10, 12) |
        prtvtept %in% c(4, 17) ~ 5,  # CDS-PP (4), PPM (17)
      
      # CHEGA
      prtvtpt == 7 |
        prtvtapt == 8 |
        prtvtbpt == 8 |
        prtvtdpt == 4 |
        prtvtept %in% c(5) ~ 6,  # CHEGA (5)
      
      # Other (inkl. Juntos pelo Povo, Partido da Terra, etc.)
      prtvtpt == 12 |
        prtvtbpt == 13 |
        prtvtcpt %in% c(11, 14, 18) | 
        prtvtdpt %in% c(21, 6) |
        prtvtept %in% c(2, 6, 8, 10, 12, 19, 20, 21, 23) ~ 9,  # Diverse kleinere Parteien
       
      # Blank/Nulo
      prtvtpt == 13 |
        prtvtapt == 12 |
        prtvtbpt == 12 |
        prtvtcpt == 17 | 
        prtvtdpt == 22 |
        prtvtept %in% c(22) ~ 10,  # Blanc/Nulo (22)
      
      # Fallback: Keep existing values
      TRUE ~ voted
    )
  )


# Recode: Voted for the Greens
ess_filtered <- ess_filtered |>
  mutate(
    vote_green = case_when(
      voted == 3 ~ 1, # Green
      voted %in% c(1, 2, 4, 5, 6, 7, 8, 9) ~ 0, # Other
      TRUE ~ NA_real_ # NA for non-voters and invalid votes
    )
  )

# Recode: Voted for the Radical Right
ess_filtered <- ess_filtered |>
  mutate(
    vote_radright = case_when(
      voted == 6 ~ 1, # Radical Right
      voted %in% c(1, 2, 3, 4, 5, 7, 8, 9) ~ 0, # Other
      TRUE ~ NA_real_ # NA for non-voters and invalid votes
    )
  )

# Define country samples



ess_filtered <- ess_filtered |>
  mutate(
    Green_sample = case_when(
      (cntry %in% c("BE", "CH", "DE", "FI", "GB", "IE", "NL", "NO", "SE")) ~ 1,
      (cntry == "FR" & year != 2023) ~ 1,  # Exclude the year 2023 for France (wave 11) -> because of NUPES not possible to look at Green voting seperately 
      TRUE ~ NA_real_
    ),
    RR_sample = case_when(
      cntry %in% c("BE", "CH", "DE", "FI", "FR", "GB", "NL", "NO", "SE") ~ 1,
      TRUE ~ NA_real_
    )
  )


df <- ess_filtered

df <- df |> 
  mutate(cohort_fct = case_when(
      cohort == 1 ~ "post-war",
      cohort == 2 ~ "60-70s",
      cohort == 3 ~ "80s",
      cohort == 4 ~ "90s",
      cohort == 5 ~ "millennial"),
    cohort_fct = cohort_fct |> 
      fct_relevel("post-war", "60-70s", "80s", "90s",
                  "millennial"),
    cohort_alt_fct = case_when(
      cohort_alt == 1 ~ "1930s",
      cohort_alt == 2 ~ "1940s",
      cohort_alt == 3 ~ "1950s",
      cohort_alt == 4 ~ "1960s",
      cohort_alt == 5 ~ "1970s",
      cohort_alt == 6 ~ "1980s",
      cohort_alt == 7 ~ "1990+"),
    cohort_alt_fct = cohort_alt_fct |> 
      fct_relevel("1930s", "1940s", "1950s", "1960s", "1970s",
                  "1980s", "1990+"),
    age_grp_fct = case_when(
      age_group == 1 ~ "18-29",
      age_group == 2 ~ "30-65",
      age_group == 3 ~ "above 65"
    ),
    edu_fct = case_when(
      education == 0 ~ "(below) lower secondary",
      education == 1 ~ "upper secondary",
      education == 2 ~ "tertiary"),
    edu_fct = edu_fct |> fct_relevel(
               "(below) lower secondary", 
               "upper secondary",
               "tertiary"),
    gndr_fct = case_when(
      gndr == 1 ~ "men",
      gndr == 2 ~ "women"),
    cntry_id_fct = factor(cntry_id),
    essround_fct = factor(essround),
    vote_abst_fct = case_when(
      voted == 0 ~ "abstention",
      voted == 1 ~ "Communist/Socialist",
      voted == 2 ~ "Social democracy",
      voted == 3 ~ "Green",
      voted == 4 ~ "Liberal",
      voted == 5 ~ "Conservative/Christian democracy",
      voted == 6 ~ "Radical right",
      voted == 7 ~ "Agrarian"),
    vote_fct = case_when(
      voted == 1 ~ "Communist/Socialist",
      voted == 2 ~ "Social democracy",
      voted == 3 ~ "Green",
      voted == 4 ~ "Liberal",
      voted == 5 ~ "Conservative/Christian democracy",
      voted == 6 ~ "Radical right",
      voted == 7 ~ "Agrarian"),
    vote_green_norr = case_match(voted,
      3 ~ 1,
      c(1,2,4,5,7) ~ 0),
    vote_rr_nogreen = case_match(voted,
      6 ~1,
      c(1,2,4,5,7) ~ 0),
    married = case_when(
      marital == 5 | maritala == 6 | maritalb == 6 ~ "never",
      marital  < 5 | maritala  < 6 | maritalb  < 6 ~ "at one point" ),
    psppsgva = ifelse(psppsgva > 5, NA, psppsgva),
    psppipla = ifelse(psppipla > 5, NA, psppipla),
    response = (psppsgva + psppipla)/2)

df <- df |> 
  dplyr::select(essround, essround_fct, cntry, cntry_id, cntry_id_fct, 
                vote_abst_fct, vote_fct, vote_green, vote_radright, 
                vote_green_norr, vote_rr_nogreen, 
                cohort, cohort_fct, cohort_alt_fct,
                age_grp_fct, class5fac, class8fac,
                education, edu_fct, edulowhigh, gndr, gndr_fct, married, response,
                agea_clean, age_group, immigrant, Green_sample, RR_sample, cntry_wave_id, pspwght, dweight, pweight)

saveRDS(df, here("data", "processed", "ess_short_1_11.rds"))

```

```{r}

```